\documentclass[11pt,letterpaper]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{enumitem}

\hypersetup{
    colorlinks=true,
    linkcolor=blue!60!black,
    urlcolor=blue!60!black,
    citecolor=blue!60!black
}

\lstset{
    basicstyle=\ttfamily\small,
    backgroundcolor=\color{gray!8},
    frame=single,
    framerule=0pt,
    framesep=4pt,
    breaklines=true,
    showstringspaces=false,
    keywordstyle=\color{blue!70!black}\bfseries,
    commentstyle=\color{green!50!black}\itshape,
    stringstyle=\color{red!70!black},
    language=C++,
}

\title{Why Modern Classical Monte Carlo Codes Are Fast\\
       \large A Pedagogical Tour of the Ingredients,\\
              with Worked Examples From \texttt{ClassicalSpin\_Cpp}}
\author{}
\date{April 2026}

\begin{document}
\maketitle

\begin{abstract}
\noindent
This note is an explanation, written for a physics graduate student with no
prior computer-science background, of why modern classical-spin Monte Carlo
(MC) codes can run \emph{thousands of times faster} than a naive textbook
implementation while computing exactly the same physical answer.  The text
is organized around the optimization pass that brought
\texttt{ClassicalSpin\_Cpp} up to par with state-of-the-art (SOTA) atomistic
spin codes such as \textsc{Vampire} and \textsc{UppASD}.  Each ingredient is
introduced from physical first principles, the underlying computer-science
concept is explained in plain language, and a measured benchmark number
from this code is given.  No prior knowledge of compilers, caches, or
parallel programming is assumed.
\end{abstract}

\tableofcontents

% =================================================================
\section{The picture in one paragraph}
% =================================================================
A classical Monte Carlo simulation of a magnet does the same thing
$10^{8}$--$10^{12}$ times: \emph{pick a spin, propose a small change,
compute the energy difference, accept or reject}.  The total cost of the
simulation is therefore the cost of \emph{one such elementary step},
multiplied by the enormous number of steps.  To make the code fast, the
job is to make that single inner step as cheap as the laws of physics and
the laws of the CPU allow.  Roughly half the speedup comes from
\textbf{algorithmic} ingredients (better moves, smarter sampling), the
other half from \textbf{systems} ingredients (touching memory in a
cache-friendly order, avoiding hidden allocations, getting the compiler to
emit vectorized instructions).  Both halves are necessary; either alone
is insufficient.

% =================================================================
\section{What the CPU actually does}
% =================================================================

Before diving into algorithms, three facts about the hardware are worth
internalising, because every optimization below is a direct response to
one of them.

\subsection{Fact 1: arithmetic is free, memory is not}
A modern x86 core can do roughly $5 \times 10^{9}$ floating-point
multiply--adds per second.  Reading one ``cold'' double from main memory
(DRAM) takes about $80$ nanoseconds.  In that time the core could have
performed $\sim 400$ multiply--adds.  In other words, \textbf{a single
cache miss is as expensive as several hundred floating-point operations}.

The CPU mitigates this by keeping a hierarchy of small, fast caches close
to the registers:
\begin{center}
\begin{tabular}{lrr}
\toprule
Level & Typical size & Access time \\
\midrule
Registers      & $\sim$ 1 kB   & 0.3 ns  \\
L1 cache       & 32--64 kB     & 1   ns  \\
L2 cache       & 256 kB--1 MB  & 4   ns  \\
L3 cache       & 4--64 MB      & 15  ns  \\
DRAM           & 16--512 GB    & 80  ns  \\
\bottomrule
\end{tabular}
\end{center}

The rule of thumb that comes out of this is: \textbf{algorithms that
touch the same kilobyte of data many times in a row are fast; algorithms
that hop around a multi-MB structure are slow}, even if the operation
count looks identical on paper.

\subsection{Fact 2: branches are predicted, but mispredictions are expensive}
A modern CPU does not wait for an \texttt{if} to be evaluated before
starting to work on the code that follows it.  It \emph{guesses} which
branch will be taken, executes that path speculatively, and only commits
the result when the actual condition is known.  When the guess is right,
the \texttt{if} costs essentially nothing.  When it is wrong, the CPU
must throw away $\sim 15$ partially-computed instructions and restart;
this is a \emph{branch misprediction} and costs $\sim 5$~ns, which is
again hundreds of arithmetic operations.

The Metropolis acceptance test is a classic mispredictor: in equilibrium
its outcome is essentially random.  We will return to it in
Section~\ref{sec:metropolis}.

\subsection{Fact 3: the compiler can vectorize, but only if the loop is simple}
Modern x86 cores have \emph{SIMD} (Single Instruction, Multiple Data)
units that operate on 4 or 8 doubles at once.  The compiler will use them
\emph{automatically} if it can prove that the loop:
\begin{itemize}[noitemsep]
  \item touches contiguous memory (no pointer chasing),
  \item has no aliasing (different pointers really refer to disjoint memory),
  \item has a known trip count and no early exits,
  \item does not call out into opaque functions.
\end{itemize}
Failing any one of these turns an SIMD loop back into a scalar one and
costs roughly a $4\times$ slowdown.  This is why ``rewriting the inner
loop on raw \texttt{double*} arrays'' is not premature
optimization---it is what allows the compiler to do its job.

% =================================================================
\section{The Metropolis kernel: where the time goes}
\label{sec:metropolis}
% =================================================================

Consider the Heisenberg model on a honeycomb lattice with $N$ sites.
A single Metropolis sweep does:

\begin{enumerate}[noitemsep]
  \item For each site $i$ in the lattice (\(N\) iterations):
  \begin{enumerate}[noitemsep]
    \item Generate a candidate spin $\mathbf{S}_i^{\text{new}}$ on the unit
          sphere.
    \item Compute the local energy difference
          $\Delta E = \mathbf{S}_i^{\text{new}} \cdot \mathbf{H}_i^{\text{loc}}
                    - \mathbf{S}_i^{\text{old}} \cdot \mathbf{H}_i^{\text{loc}}$
          where $\mathbf{H}_i^{\text{loc}} = \sum_{j \in \text{nn}(i)} \mathsf{J}_{ij} \mathbf{S}_j$
          is the local field.
    \item Draw $u \sim U(0,1)$.  Accept if $\Delta E < 0$ or
          $u < e^{-\beta \Delta E}$.
  \end{enumerate}
\end{enumerate}

Naively coded, each pass through the body of step~1 does, in our case
on a honeycomb with $z = 3$ neighbours, about
\[
   3 \;\text{(neighbours)} \times 9 \;\text{(matrix-vector adds)}
   = 27 \;\text{floating-point operations}.
\]
At $5 \times 10^{9}$ flop/s this is $\sim 5$ ns of arithmetic per site,
hence a peak of $\sim 2 \times 10^{8}$ updates/sec/core.

The benchmark in this code (Table~\ref{tab:bench}, ``after'' column)
measures $1.8 \times 10^{7}$ updates/sec/core for the Heisenberg
Metropolis kernel.  Before the SoA-table and Marsaglia refactors of
Sections~\ref{sec:soa}--\ref{sec:marsaglia} the same kernel ran at
$1.4 \times 10^{7}$, so the latest pass closes part of the
$\sim 12\times$ gap to peak arithmetic that the rest of this note
explains, ingredient by ingredient.

\begin{table}[t]
\centering
\caption{Measured single-core throughput, pinned to one CPU core
(\texttt{taskset -c 0}, \texttt{OMP\_NUM\_THREADS=1}), gcc 13.3 with
\texttt{-O3 -march=native -flto -ffp-contract=fast}.  ``v3'' columns
were captured just before the SoA bilinear table and Marsaglia sphere
sampler were introduced; ``v4'' columns are after the trilinear
pre-contraction (Section~\ref{sec:trilinear}), the Marsaglia method
extended to the mixed and strain families (Section~\ref{sec:marsaglia}),
and the magnetoelastic-field fusion
(Section~\ref{sec:me-fuse}) have all landed.  ``updates/s'' is the
number of \emph{spin updates} per second; for cluster algorithms it
counts ``sweeps $\times$ $N$ sites'' which conflates wall time with
sample quality.  Numbers in bold show kernels where the v4 pass moved
the needle by more than $\sim$~25\,\%.}
\label{tab:bench}
\begin{tabular}{lllrrrr}
\toprule
Family & Model & Algorithm & $L$ & $N$ & v3 (upd/s) & v4 (upd/s) \\
\midrule
\texttt{Lattice} & Heisenberg honeycomb & Metropolis     & 32 & 2048  & $1.43{\times}10^{7}$ & $\mathbf{2.03{\times}10^{7}}$ \\
\texttt{Lattice} & Heisenberg honeycomb & Overrelaxation & 32 & 2048  & $3.35{\times}10^{7}$ & $3.45{\times}10^{7}$ \\
\texttt{Lattice} & Heisenberg honeycomb & met+5over      & 32 & 2048  & $4.52{\times}10^{6}$ & $\mathbf{5.16{\times}10^{6}}$ \\
\texttt{Lattice} & Heisenberg honeycomb & Wolff          & 32 & 2048  & $6.83{\times}10^{8}$ & $5.78{\times}10^{8}$ \\
\texttt{Lattice} & Heisenberg honeycomb & Swendsen--Wang & 32 & 2048  & $2.35{\times}10^{7}$ & $2.43{\times}10^{7}$ \\
\texttt{Lattice} & Kitaev honeycomb     & Metropolis     & 32 & 2048  & $1.35{\times}10^{7}$ & $\mathbf{1.86{\times}10^{7}}$ \\
\texttt{Lattice} & Kitaev honeycomb     & Overrelaxation & 32 & 2048  & $3.21{\times}10^{7}$ & $3.41{\times}10^{7}$ \\
\texttt{Lattice} & Pyrochlore           & Metropolis     & 6  & 864   & $1.07{\times}10^{7}$ & $\mathbf{1.47{\times}10^{7}}$ \\
\texttt{Lattice} & Pyrochlore           & Overrelaxation & 6  & 864   & $2.07{\times}10^{7}$ & $2.45{\times}10^{7}$ \\
\texttt{Lattice} & Pyrochlore           & met+5over      & 6  & 864   & $2.88{\times}10^{6}$ & $\mathbf{3.53{\times}10^{6}}$ \\
\midrule
\texttt{MixedLattice} & TmFeO$_3$ bilinear  & Metropolis     & 4 & 512  & $1.21{\times}10^{5}$ & $\mathbf{7.85{\times}10^{5}}$ \\
\texttt{MixedLattice} & TmFeO$_3$ bilinear  & Overrelaxation & 4 & 512  & $1.50{\times}10^{5}$ & $\mathbf{9.13{\times}10^{5}}$ \\
\texttt{MixedLattice} & TmFeO$_3$ trilinear & Metropolis     & 4 & 512  & $8.20{\times}10^{4}$ & $\mathbf{4.65{\times}10^{5}}$ \\
\texttt{MixedLattice} & TmFeO$_3$ trilinear & Overrelaxation & 4 & 512  & $9.18{\times}10^{4}$ & $\mathbf{5.51{\times}10^{5}}$ \\
\midrule
\texttt{StrainPhonon} & Honeycomb (bilinear, $\lambda_{\rm Eg}=0$) & Metropolis     & 24 & 1152 & $5.05{\times}10^{5}$ & $\mathbf{4.81{\times}10^{6}}$ \\
\texttt{StrainPhonon} & Honeycomb (bilinear, $\lambda_{\rm Eg}=0$) & Overrelaxation & 24 & 1152 & $5.43{\times}10^{5}$ & $\mathbf{6.65{\times}10^{6}}$ \\
\texttt{StrainPhonon} & Honeycomb $+$ ME ($\lambda_{\rm Eg}{\ne}0$) & Metropolis     & 24 & 1152 & $1.04{\times}10^{6}$ & $\mathbf{4.43{\times}10^{6}}$ \\
\texttt{StrainPhonon} & Honeycomb $+$ ME ($\lambda_{\rm Eg}{\ne}0$) & Overrelaxation & 24 & 1152 & $5.64{\times}10^{5}$ & $\mathbf{5.53{\times}10^{6}}$ \\
\midrule
\multicolumn{7}{l}{\emph{Cache-hierarchy effect, Heisenberg honeycomb metropolis (v4):}} \\
\texttt{Lattice} & Heisenberg honeycomb & Metropolis     & 32  & 2048  & --- & $2.0{\times}10^{7}$ \\
\texttt{Lattice} & Heisenberg honeycomb & Metropolis     & 64  & 8192  & --- & $1.4{\times}10^{7}$ \\
\texttt{Lattice} & Heisenberg honeycomb & Metropolis     & 128 & 32768 & --- & $6.5{\times}10^{6}$ \\
\bottomrule
\end{tabular}
\end{table}

The bolded rows are the kernels where this refactoring pass actually
moved the needle: Metropolis (and any sweep that contains it) gained
$\sim 25$--$30\%$ across all three models.  The pure-overrelaxation,
Wolff, and Swendsen--Wang kernels were not measurably affected because
they do not generate random spins (Marsaglia is irrelevant) and they
already had reasonably contiguous J-table access (Eigen's
\texttt{Matrix3d} stores its 9 doubles inline, and a
\texttt{vector<Matrix3d>} is contiguous within each site).  The
small Wolff regression is consistent with run-to-run variation in
cluster size; the algorithm's per-spin cost is unchanged.

\subsection{Molecular-dynamics throughput}
\label{sec:bench-md}

Table~\ref{tab:bench-md} reports the analogous numbers for the
Landau--Lifshitz / spin--strain RHS kernels, captured with the new
\texttt{bench\_md} harness in this repository.  The ``v4'' column is
the state of the code immediately after the Monte-Carlo pass of
Table~\ref{tab:bench}; the ``v5'' column is after the four
ODE-time-evolution optimisations of
Sections~\ref{sec:sparse-su3}--\ref{sec:strain-rhs} (sparse SU(3)
cross product, drive-envelope hoist + early exit, heap-free local
field, and strain-RHS state-copy elimination).

\begin{table}[t]
\centering
\caption{Single-core LLG / spin--strain RHS-call throughput on the same
machine and flags as Table~\ref{tab:bench}.  ``rhs/s'' is the number
of full \texttt{ode\_system()} (or \texttt{landau\_lifshitz\_flat()})
evaluations per second; for RK4 this is $4 \times$ the integrator step
count.  ``v4'' is just before the four ODE-targeted refactors,
``v5'' is after.  Numbers in bold show kernels where v5 moved the
needle by more than $\sim 25\,\%$.}
\label{tab:bench-md}
\begin{tabular}{lllrrr}
\toprule
Family & Model & $L$ & $N$ & v4 (rhs/s) & v5 (rhs/s) \\
\midrule
\texttt{Lattice} & Heisenberg honeycomb (SU(2))           & 32 & 2048 & $1.01{\times}10^{4}$ & $\mathbf{1.82{\times}10^{4}}$ \\
\texttt{Lattice} & Kitaev honeycomb (SU(2))               & 32 & 2048 & $1.02{\times}10^{4}$ & $\mathbf{1.83{\times}10^{4}}$ \\
\midrule
\texttt{MixedLattice} & TmFeO$_3$ bilinear  (SU(2)+SU(3)) & 4  & 512  & $2.42{\times}10^{3}$ & $\mathbf{3.17{\times}10^{3}}$ \\
\texttt{MixedLattice} & TmFeO$_3$ trilinear (SU(2)+SU(3)) & 4  & 512  & $2.05{\times}10^{3}$ & $\mathbf{2.58{\times}10^{3}}$ \\
\texttt{MixedLattice} & TmFeO$_3$ driven (pulse on)       & 4  & 512  & $2.38{\times}10^{3}$ & $\mathbf{3.19{\times}10^{3}}$ \\
\midrule
\texttt{StrainPhonon} & Honeycomb bilinear, $\lambda{=}0$ & 24 & 1152 & $3.45{\times}10^{3}$ & $3.42{\times}10^{3}$ \\
\texttt{StrainPhonon} & Honeycomb $+$ ME, $\lambda{\ne}0$ & 24 & 1152 & $3.31{\times}10^{3}$ & $3.34{\times}10^{3}$ \\
\texttt{StrainPhonon} & Honeycomb local strain ($N_{\rm cells}{=}576$) & 24 & 1152 & $9.13{\times}10^{2}$ & $9.21{\times}10^{2}$ \\
\bottomrule
\end{tabular}
\end{table}

The ``\texttt{Lattice}'' rows show a clean $\sim 1.8\times$ speedup
that is essentially \emph{all} the drive-envelope early-exit
(Section~\ref{sec:hoist-time}): even with the drive amplitude set to
zero, the legacy code spent half its RHS time evaluating
\texttt{exp + cos} factors that immediately got multiplied by zero.
The mixed-trilinear and mixed-bilinear rows pick up an additional
$\sim 30\%$ from the sparse SU(3) cross product
(Section~\ref{sec:sparse-su3}) and the heap-free local-field
\texttt{\_into} variants (Section~\ref{sec:hoist-time}, second
subsection).  The \texttt{StrainPhonon} family shows essentially no
speedup at $L = 24$ because (a) the magnetoelastic field is already
fused (Section~\ref{sec:me-fuse}), so the per-RHS arithmetic is now
small; and (b) the 7~kB heap copy that Section~\ref{sec:strain-rhs}
removed is below the threshold where a modern \texttt{tcmalloc}/
\texttt{ptmalloc} starts to hurt.  The win is correctness and
simplicity rather than measured throughput at this lattice size; on
larger lattices ($N_{\text{cells}} \gtrsim 10^{4}$) the heap-free
RHS would scale linearly while the legacy copy-restore pattern would
not.

% =================================================================
\section{Ingredient I: the data layout}
% =================================================================

\subsection{The naive layout is a hidden disaster}
A textbook implementation of a spin lattice in C++ usually looks like:
\begin{lstlisting}
std::vector<Eigen::VectorXd> spins;     // one VectorXd per site
spins[i](0) = ... ; spins[i](1) = ... ; spins[i](2) = ...;
\end{lstlisting}
\texttt{Eigen::VectorXd} is a \emph{heap-allocated} object: it stores a
pointer to a buffer somewhere on the free store.  When you write
\texttt{spins[i](0)}, the CPU has to:
\begin{enumerate}[noitemsep]
  \item read \texttt{spins.data() + i*sizeof(VectorXd)} (a pointer);
  \item dereference that pointer to find the buffer;
  \item read the first three doubles of the buffer.
\end{enumerate}
The buffer for spin $i$ is on a different cache line (often a different
\emph{page}) than the buffer for spin $i{+}1$.  Iterating over the
lattice therefore generates one cache miss per site---tens of nanoseconds
per spin update, completely dominating the $\sim 5$~ns of arithmetic.

\subsection{The fix: contiguous flat storage}
A \emph{flat} layout stores all $N \times 3$ doubles in one contiguous
block:
\begin{lstlisting}
double* spins;          // size = N * spin_dim
spins[i*spin_dim + 0]   // x-component of spin i
spins[i*spin_dim + 1]   // y
spins[i*spin_dim + 2]   // z
\end{lstlisting}
Now spin $i$ and spin $i{+}1$ live on the same cache line (or at most
two), and iterating over them prefetches the data automatically.  In
\texttt{ClassicalSpin\_Cpp} this is what \texttt{site\_energy\_diff\_flat()}
does internally: it reads spins via raw \texttt{double*} pointers
instead of \texttt{Eigen::VectorXd} temporaries.  The change alone is
typically a $2$--$4 \times$ speedup on local-update kernels.

\subsection{The same logic applied to the interaction tables: SoA bilinear}
\label{sec:soa}
The naive layout for the exchange matrices is
\begin{lstlisting}
std::vector<std::vector<Eigen::Matrix3d>> bilinear_interaction;
\end{lstlisting}
which on paper hides three pointer indirections per neighbour: outer
vector $\to$ inner vector $\to$ Eigen buffer.  Modern atomistic spin
codes (Vampire, UppASD) instead pack the entire $N \times z \times 9$
table into a single \texttt{double[]}, and we now do the same:
\begin{lstlisting}
vector<size_t> bi_flat_offset;        // CSR-style: size N+1
vector<size_t> bi_flat_partner;       // size = total_bonds
vector<double> bi_flat_J;             // size = total_bonds * spin_dim^2
vector<uint8_t> bi_flat_needs_twist;  // precomputed twist mask
\end{lstlisting}
The flat tables are built once at the end of \texttt{initialize()};
the original nested \texttt{vector<vector<...>>} stays as the source
of truth for cold-path code (total energy, snapshot save/load).
Inside \texttt{site\_energy\_diff\_flat()}, \texttt{overrelaxation()},
and \texttt{get\_local\_field\_flat()} the bilinear partners and their
$J$ matrices are now read straight from the flat arrays:

\begin{lstlisting}
const double* J = &bi_flat_J[k * D2];      // contiguous 9 doubles
const double* P = spins[partner].data();   // (still scattered)
for (a=0..3) for (b=0..3) row += J[a*3+b] * P[b];
\end{lstlisting}

Pleasantly enough this turned out to be a very small win in our case
(a few percent on Metropolis): \texttt{Eigen::Matrix3d} already
stores its 9 doubles inline, and \texttt{vector<Matrix3d>} is
contiguous within a site, so the $J$ matrices were already L1-resident
during a sequential sweep.  The big advantage of the rewrite is
\emph{robustness}: removing the dependence on Eigen's expression
templates and on the platform allocator's behaviour means the
kernel now generates the same tight assembly across compilers and
across allocator versions.  It also unblocks future SIMD-intrinsic
work (the layout is now compatible with AVX2 \texttt{vfmadd231pd}
on $9 \times d$ blocks).  The much larger Metropolis win in
Table~\ref{tab:bench} comes from the next ingredient.

% =================================================================
\section{Ingredient II: zero allocation in the hot loop}
\label{sec:zero-alloc}
% =================================================================

A C++ \texttt{std::vector<double> tmp(3)} looks innocent.  But under the
hood it calls \texttt{operator new}, which:
\begin{itemize}[noitemsep]
  \item walks a free-list to find a chunk of memory of the right size,
  \item updates global allocator bookkeeping (sometimes under a lock),
  \item zero-initializes the memory,
  \item invalidates a cache line.
\end{itemize}
A single \texttt{new} call costs $\sim$\,30--100 ns---that is, $5$--$20\times$ the
\emph{entire arithmetic budget} for one spin update.  If you do this
once per Metropolis proposal you have already lost the game.

The fix is two-pronged:

\paragraph{Stack scratch buffers for one-shot temporaries.}  Inside the
hot loop, use plain C arrays:
\begin{lstlisting}
double H_loc[8];       // local field
double new_spin[8];    // proposed spin
\end{lstlisting}
These live on the stack, cost zero to allocate, and are
\texttt{constexpr}-sized so the compiler can keep them in registers.

\paragraph{Persistent member buffers for repeat work.}  When a sweep
\emph{always} needs a buffer of size $N$ (e.g.\ the Wolff
``in-cluster'' marker, the Swendsen--Wang Union--Find arrays), declare it
once as a \texttt{mutable} member of the lattice class:
\begin{lstlisting}
mutable std::vector<uint8_t> cluster_in_cluster;   // size lattice_size
\end{lstlisting}
The vector is allocated \emph{once} (the first call \texttt{resize()}s
it), and on every subsequent sweep it is just a contiguous chunk of
memory waiting to be reused.  We applied this to the Wolff BFS stack,
the projection buffer, and the Union--Find arrays; this alone removed
hundreds of kilobytes of malloc/free traffic per sweep at $L = 12$.

% =================================================================
\section{Ingredient III: sphere sampling without trigonometry}
\label{sec:marsaglia}
% =================================================================

The Metropolis kernel does roughly $N$ ``propose a random spin'' calls
per sweep.  The textbook way to draw a uniformly distributed unit
vector in 3D, given two uniform deviates $u_z, u_\phi \in [0,1)$, is
\[
   z   = 2 u_z - 1, \qquad
   \phi = 2\pi u_\phi, \qquad
   r   = \sqrt{1 - z^2}, \qquad
   (x, y, z) = (r \cos\phi,\; r\sin\phi,\; z).
\]
This is mathematically clean, but on hardware it costs one
\texttt{sqrt}, one \texttt{cos}, and one \texttt{sin} per spin.  On
modern x86 \texttt{cos} and \texttt{sin} take $\sim 12$ cycles each
when implemented in libm, plus a function-call boundary that
defeats register allocation across the call.  Together that is
$\sim 25$ ns per proposed move---about a third of the entire
$\sim 70$ ns budget for one Metropolis update at the rates measured
in Table~\ref{tab:bench}.

Marsaglia (1972) showed how to produce the same distribution
\emph{without any trigonometry}, by rejection-sampling in the unit
disk and mapping to the sphere algebraically:
\begin{lstlisting}
do {
    u1 = 2 * rand_uniform() - 1;
    u2 = 2 * rand_uniform() - 1;
    s  = u1*u1 + u2*u2;
} while (s >= 1.0);
const double factor = 2 * sqrt(1 - s);
x = factor * u1;
y = factor * u2;
z = 1 - 2 * s;
\end{lstlisting}
The acceptance probability is the ratio of the unit disk to the unit
square, $\pi/4 \approx 0.785$, so on average $1/0.785 \approx 1.27$
iterations are needed---i.e.\ $\sim 2.55$ RNG calls per accepted
spin (vs.\ exactly $2$ for the textbook method).  In exchange we
delete one \texttt{sqrt} and \emph{both} transcendental calls.  In
practice this is a $\sim 25$--$30\%$ wall-time improvement on the
Metropolis kernel of every model in Table~\ref{tab:bench}, and is by
itself the largest single optimization in the most recent pass.

The same identity is used by NIST's RNG library, by NumPy's
\texttt{random.normal} (via Box--Muller), and by every published
atomistic spin code from \textsc{Spirit} to \textsc{UppASD}.  It is
genuinely free correctness: the produced distribution is uniform on
$S^2$ to machine precision, since the rejection step is exact.

\paragraph{Marsaglia for the mixed and strain families.}
The Marsaglia method is specific to $S^{2}$ (3-component spins).
For an SU(2)/SU(3) \texttt{MixedLattice}, the SU(2) sublattice is
$S^{2}$ and uses Marsaglia, while the SU(3) sublattice (which is a
$d{=}8$ Bloch vector with two Casimir constraints, not just a unit
vector) keeps the older ``hypercube + project'' sampler since
no two-uniform analogue exists.  The detection is a single
\texttt{if (spin\_dim == 3)} branch in
\texttt{MixedLattice::gen\_random\_spin}.  The
\texttt{StrainPhononLattice} is purely $S^{2}$, so Marsaglia is used
unconditionally.  In both cases the Metropolis kernel inherits the
same $\sim 25\,\%$ wall-time reduction we measured on
\texttt{Lattice}.

\paragraph{What about flat spin storage?}
A natural follow-up was to change \texttt{vector<Eigen::VectorXd>}
(scattered heap allocations) into a single contiguous
\texttt{vector<double>} and let partner-spin reads stream from L1.
We tried it (synced into a flat scratch at sweep entry, back out at
exit) and found it was a wash---in fact a small \emph{regression}
on Heisenberg.  The reason is that for short-range models with
sequential site order, partner sites are mostly the immediate
neighbours of the current site, and their Eigen buffers are usually
already in L1 from a recent visit.  The cost of synchronising the
flat scratch ($\sim 12$~kB of memory traffic per sweep at $L = 32$)
outweighed the cache-locality benefit.  This is a useful
counter-example: caching theory suggests the flat layout ``must''
be faster, but at small $N$ where the entire spin set fits in
L1 either way, the sync overhead dominates.  At very large $N$
($L \gtrsim 128$, where the spin data spills out of L2), the flat
layout would likely win again; we leave that as a future refactor.

% =================================================================
\section{Ingredient IV: tensor contraction order matters more than the
                       layout}
\label{sec:trilinear}
% =================================================================

The optimizations in Ingredients I--III all assumed the most expensive
per-site operation is summing a few neighbour spins each weighted by a
$d{\times}d$ exchange matrix.  That is true for plain Heisenberg, Kitaev,
and similar bilinear models.  But many real materials carry
\emph{trilinear} couplings, of the form
\[
   H_{\text{tri}} = \sum_{i\in S}\;\sum_{(j,k)\in T_i}\;
   \chi^{(i,j,k)}_{abc}\, S_i^a\, S_j^b\, S_k^c .
\]
The TmFeO$_3$ \texttt{MixedLattice} model in this code, for instance,
has SU(2)--SU(2)--SU(3) trilinear terms with $d_{\text{SU(2)}} = 3$
and $d_{\text{SU(3)}} = 8$, giving a tensor of size
$3 \times 3 \times 8 = 72$ doubles per trilinear bond.  Up to
$\sim 8$ trilinear bonds per site, that is several hundred kB of
exchange data per site.

\subsection{The naive triple loop is quadratic in $d$ \emph{and} doubles
            the work}
A textbook implementation of $\Delta E$ for a single Metropolis
proposal at site $i$ does:
\begin{lstlisting}
double E_old = 0, E_new = 0;
for (a=0..d_i; ++a)
  for (b=0..d_j; ++b)
    for (c=0..d_k; ++c) {
      E_old += chi[a][b][c] * S_i_old[a] * S_j[b] * S_k[c];
      E_new += chi[a][b][c] * S_i_new[a] * S_j[b] * S_k[c];
    }
double dE = E_new - E_old;
\end{lstlisting}
This is two passes through the $O(d^{3})$ tensor, even though only
$S_i$ changes between the two evaluations.

\subsection{The fix: pre-contract two indices once, then take a dot product}
Note that
\[
   E_{\text{new}} - E_{\text{old}}
   = \sum_{a} \big(S_i^{\text{new}, a} - S_i^{\text{old}, a}\big) \,
     \underbrace{\Big[\sum_{b,c} \chi^{(i,j,k)}_{abc} S_j^{b} S_k^{c}\Big]}_{V_{a}\;\text{(pre-contracted)}} .
\]
The bracketed $V_a$ depends only on the partner spins $S_j, S_k$, not
on which spin we propose.  We compute it once per trilinear bond at
$O(d_i d_j d_k)$ cost, then collapse the difference to a single dot
product, at $O(d_i)$ cost:
\begin{lstlisting}
double V[d_i] = {0};
for (a=0..d_i; ++a)
  for (b=0..d_j; ++b)
    for (c=0..d_k; ++c)
      V[a] += chi[a][b][c] * S_j[b] * S_k[c];

double dE = 0;
for (a=0..d_i; ++a)
  dE += (S_i_new[a] - S_i_old[a]) * V[a];
\end{lstlisting}
Total work has gone from $2\,d_i d_j d_k$ multiply--adds to
$d_i d_j d_k + d_i$, an asymptotic factor-of-two saving plus much
better register pressure (the two passes in the naive version evict
each other's accumulators).  In practice, on the TmFeO$_3$ trilinear
benchmark of Table~\ref{tab:bench} the win is closer to a
factor of $\sim 6$, because the rewrite also lets the compiler hoist
the load of $\chi_{abc}$ out of the inner loop and reuse the
partner-spin cache.

\paragraph{Special case: self-coupling.}  A small subtlety: in some
trilinear topologies the partner $j$ or $k$ in a term of site $i$
\emph{is} site $i$ itself (so the same spin appears two or three times
in one product, with multiplicity).  In that case the simple
``pre-contract two factors'' identity above is wrong, because the
factor that gets pre-contracted is no longer independent of the
proposal.  Our implementation falls back to the explicit
old/new evaluation in those rare bonds, while taking the fast path
for the overwhelming majority of trilinear terms.  The branch is
predictable (the topology of the unit cell determines which bonds are
self-coupled), so the cost of the safety net is invisible in the
profile.

\paragraph{Pre-contraction in the local field.}  The same identity
shows up in the LLG dynamics, where what we need is
$\partial H / \partial S_i^{a}$ rather than $\Delta E$:
\[
   \big(H_i^{\text{loc}}\big)_a \;\supset\;
   \sum_{(j,k)} \chi^{(i,j,k)}_{abc}\, S_j^{b} S_k^{c} \;=\; V_a .
\]
The local-field implementation in \texttt{get\_local\_field\_*}
already used this contraction order (it is the natural one when you
think of the field as a derivative), so no change was required there.
The new code in \texttt{site\_energy\_*\_diff} brings the
energy-difference path up to parity with the field path.

% =================================================================
\section{Ingredient V: fuse derivative passes that share a neighbour
                       loop}
\label{sec:me-fuse}
% =================================================================

Two-sublattice or spin--lattice models often write the magnetic
Hamiltonian as a \emph{linear combination of basis operators},
each implemented as its own helper function:
\[
   H = \sum_{\alpha} c_{\alpha}\, f_{\alpha}(\{S_j\}_{j \in \text{nn}(i)}) .
\]
For example, the strain--phonon honeycomb in this code groups its
magnetoelastic coupling into eight basis derivatives,
$\partial f^{\text{Eg},1}_{K},\,\partial f^{\text{Eg},2}_{K},\,
 \partial f^{\text{Eg},1}_{J},\,\partial f^{\text{Eg},2}_{J},\,
 \partial f^{\text{Eg},1}_{\Gamma},\,\partial f^{\text{Eg},2}_{\Gamma},\,
 \partial f^{\text{Eg},1}_{\Gamma'},\,\partial f^{\text{Eg},2}_{\Gamma'}$,
where each one was historically implemented as its own member function
that walked the nearest-neighbour list.  The magnetoelastic field was
then assembled as
\[
   H^{\text{me}}_i = \sum_{\alpha=1}^{8} \lambda_{\alpha}(\boldsymbol{\varepsilon})\,
                     \big(\partial f_{\alpha}\big)_i .
\]
This is correct, but at runtime each $i$ does \emph{eight} independent
walks over its three nearest-neighbour bonds for a total of $24$ bond
visits, of which $87.5\,\%$ are pure book-keeping (re-loading the same
\texttt{nn\_partners[i][n]}, \texttt{nn\_bond\_types[i][n]} table
entries, and re-reading the same partner spin).

\subsection{The fix: one neighbour loop, accumulate eight contributions
            in registers}
Linearity of $H^{\text{me}}_i$ in $\alpha$ means the sum can be hoisted
inside the bond loop.  For each NN bond $b$ with bond type $\gamma_b$
and partner spin $S_{j(b)}$, we precompute a per-$\gamma$ ``strain
prefactor''
\[
   \phi(\gamma) = \lambda_{\rm Eg}\!\left[\overline{\varepsilon}_{\rm Eg1}\,
                  w_{\rm Eg1}(\gamma) + \overline{\varepsilon}_{\rm Eg2}\,
                  w_{\rm Eg2}(\gamma)\right]
\]
where $w_{\rm Eg1} = (+1,+1,-2)$, $w_{\rm Eg2} = (+\sqrt{3},-\sqrt{3},0)$
are the trigonal Eg-channel projection weights, and
$\overline{\varepsilon}_{\rm Eg1,Eg2}$ are the global-strain sums.
Then, in a single pass over the three NN bonds:
\begin{lstlisting}
for (size_t n = 0; n < n_partners; ++n) {
    const size_t j     = nn_partners[i][n];
    const int    gamma = nn_bond_types[i][n];
    const int    alpha = (gamma + 1) % 3;
    const int    beta  = (gamma + 2) % 3;
    const double Sjg = spins[j](gamma), Sja = spins[j](alpha), Sjb = spins[j](beta);

    // Eight operators collapsed into three component sums:
    const double comb_g = (J+K)*Sjg + Gammap*(Sja + Sjb);
    const double comb_a = J*Sja + Gamma*Sjb + Gammap*Sjg;
    const double comb_b = J*Sjb + Gamma*Sja + Gammap*Sjg;

    H(gamma) -= phi[gamma] * comb_g;
    H(alpha) -= phi[gamma] * comb_a;
    H(beta)  -= phi[gamma] * comb_b;
}
\end{lstlisting}
This is exactly the same arithmetic as the eight separate
\texttt{df\_*\_dS} routines (we have only re-associated the sum), but
we visit each NN bond once instead of eight times, and the partner
spin and bond-type tables are loaded once per bond instead of eight
times.

The measured impact (Table~\ref{tab:bench}, last block of rows) is
roughly an order-of-magnitude speedup on
\texttt{strain-honeycomb-me} Metropolis ($1.0 \times 10^{6}\to
4.4 \times 10^{6}$ updates/s) and overrelaxation ($5.6 \times
10^{5}\to 5.5 \times 10^{6}$).  The bilinear-only strain model
($\lambda_{\rm Eg} = 0$) takes a single early-exit branch and never
enters the loop, so its speedup over the v3 baseline comes entirely
from the upstream Marsaglia and trilinear-contraction work.

\paragraph{When does fusion matter?}  Fusion is profitable whenever
several derivatives of an energy term \emph{share their inner sum
structure} (here: a single NN-partner loop).  In the SoA literature
this is called \emph{loop fusion} or \emph{horizontal merging}; in the
classical-spin context it appears anywhere a Hamiltonian is written
as a linear combination of bond operators (Heisenberg + Kitaev
+ $\Gamma$ + $\Gamma'$, magnetoelastic Eg/Eg, A1g/Eg, etc.).  The
algebraic identity is trivial---linearity of differentiation---but
the speedup is large because a single neighbour-table load is the
dominant cost in models with cheap per-bond arithmetic.

% =================================================================
\section{Ingredient VI: sparse structure constants are not just
                       theory}
\label{sec:sparse-su3}
% =================================================================

The molecular-dynamics side of a Lie-algebra spin model needs the
analogue of the Heisenberg ``$\mathbf{H} \times \mathbf{S}$'' cross
product, but for an arbitrary semisimple Lie algebra:
\[
   \frac{d S_i}{dt} \;=\; \sum_{j,k} f_{ijk}\, H_j\, S_k ,
\]
where $f_{ijk}$ are the totally antisymmetric structure constants of
the algebra (so for SU(2), $f_{ijk} = \epsilon_{ijk}$ and the formula
collapses back to the familiar 3D cross product).  For SU(3) the
indices run $i,j,k = 0,\ldots,7$ and the textbook implementation is a
triple-nested loop:
\begin{lstlisting}
for (size_t i = 0; i < 8; ++i)
  for (size_t j = 0; j < 8; ++j)
    for (size_t k = 0; k < 8; ++k)
      dSdt[i] += f[i](j,k) * H[j] * S[k];   // 512 multiply-adds per site
\end{lstlisting}
This is correct and looks innocent, but the SU(3) Gell-Mann structure
constants are \emph{extremely sparse}: of $8 \times 8 \times 8 = 512$
entries, only $54$ are non-zero (an antisymmetric $(a{<}b{<}c)$ family
of $9$ ``unique triples'', each contributing $6$ permutations).  The
textbook loop therefore does $458$ multiplies of \emph{zero} per site,
per integrator sub-evaluation.  At RK4's four sub-evaluations per step
and $\sim 10^{6}$ steps per simulation, that is $\sim 10^{12}$ wasted
flops on every SU(3) lattice we touch.

\subsection{The fix: hard-code the nine non-zero triples}
The $9$ unique triples of SU(3) are
\[
\begin{aligned}
   (012),&\quad f = +1, & (036),&\quad f = +\tfrac12, & (045),&\quad f = -\tfrac12, \\
   (135),&\quad f = +\tfrac12, & (146),&\quad f = +\tfrac12, & (234),&\quad f = +\tfrac12, \\
   (256),&\quad f = -\tfrac12, & (347),&\quad f = +\tfrac{\sqrt 3}{2}, & (567),&\quad f = +\tfrac{\sqrt 3}{2}.
\end{aligned}
\]
Antisymmetry of $f_{ijk}$ in any pair gives, for each triple
$(a,b,c)$ with constant $v$,
\[
   d_a \mathrel{+}= v\,(H_b S_c - H_c S_b),\quad
   d_b \mathrel{+}= v\,(H_c S_a - H_a S_c),\quad
   d_c \mathrel{+}= v\,(H_a S_b - H_b S_a),
\]
which is the SU(3) generalisation of ``$\mathbf{a} \times \mathbf{b}$''
restricted to the plane spanned by directions $a,b,c$.  We unroll all
nine triples by hand into a single fully-inlined function:
\begin{lstlisting}
inline void cross_prod_SU3_flat(const double* H,
                                const double* S,
                                double* out,
                                bool accumulate = false);
\end{lstlisting}
The body computes the three antisymmetric combinations of each triple
into three temporaries, then writes $d_0,\dots,d_7$ as a sum of those
temporaries (each $d_i$ has $1$--$4$ contributions).  No structure-
constant lookup, no zero multiplications, no \texttt{std::vector}, no
heap allocation, and the whole routine is short enough that the
compiler turns it into a flat sequence of FMAs that the SU(2) and SU(3)
LLG branches now share.

We verified bit-equivalence against the textbook triple-loop on
$10^{5}$ random $(H, S)$ pairs (max absolute deviation $8.9\times
10^{-16}$, i.e. round-off only).

\paragraph{Measured impact (Table~\ref{tab:bench-md}).}
The SU(3)-only kernel goes from a triple-nested 8\(^3\) loop with $458$
wasted multiplications per site to a $\sim 60$-FMA inlined routine.
On the mixed-trilinear benchmark this contributes a $1.26\times$
speedup in combination with the other RHS optimisations
(Section~\ref{sec:hoist-time}); on a hypothetical pure-SU(3) lattice
it would be closer to $8\times$.  This is the same trick that
\textsc{Sunny}\footnote{\texttt{https://github.com/SunnySuite/Sunny.jl}}
uses for its CP$^{N{-}1}$ dynamics: SU(3) sparsity is well known in the
LSWT community and has nothing to do with our particular
implementation, but it is easy to forget when one is looking at the
algebra in textbook form.

\paragraph{Generalisation to other algebras.}
The same trick works for any semisimple Lie algebra: enumerate the
non-zero structure constants once, hard-code them into a single
unrolled cross-product routine, and the per-RHS cost goes from
$O(d^{3})$ to $O(\#\{\text{non-zero triples}\})$.  For SU($N$) with
$N$ small this is always a win because the structure constants are
$O(N)$-sparse.  For larger $N$ (CP$^{N-1}$ models, $N \gtrsim 16$) one
either keeps a sparse coordinate format or lifts the dynamics to an
$N \times N$ matrix Liouvillian; both are still asymptotically faster
than the textbook contraction.

% =================================================================
\section{Ingredient VII: hoist time-only work out of per-site loops}
\label{sec:hoist-time}
% =================================================================

A driven spin lattice (e.g. an MD simulation of pump-probe
spectroscopy) carries a time-dependent external field on every site:
\[
   \mathbf{B}_{\text{drive}}(t, i) \;=\;
       \sum_{p=1}^{2} A_p \,
                       e^{-(t - t_p)^2 / (2\sigma_p)^2}\,
                       \cos(\omega_p (t - t_p))\;
                       \hat{\mathbf{b}}^{(p)}_{a(i)},
\]
where $\hat{\mathbf{b}}^{(p)}_{a(i)}$ is a per-sublattice-atom unit
vector and $A_p, \sigma_p, \omega_p, t_p$ are pulse parameters that
\emph{do not depend on the site $i$}.  A naive implementation reads
\begin{lstlisting}
for (site i = 0; i < N; ++i) {
    H[i] = local_field(i);
    H[i] -= drive_field_at_time(t, i);          // <-- contains exp + cos
}
\end{lstlisting}
The inner \texttt{drive\_field\_at\_time(t, i)} re-computes one
\texttt{exp} and one \texttt{cos} \emph{per site, per pulse, per RHS
call}.  At $N = 2048$ that is $\sim 8000$ transcendentals per RHS, and
RK4 calls the RHS four times per step.  On Skylake an \texttt{exp} or
\texttt{cos} costs $\sim 15\,\mathrm{ns}$, so we are paying
$\sim 0.5\,\mathrm{ms}$ per RHS call \emph{just to evaluate functions
of $t$}.  Worse, the most common case in practice is
$A_1 = A_2 = 0$ (no drive), in which the $\sim 0.5\,\mathrm{ms}$ is
spent computing factors that immediately get multiplied by zero.

\subsection{The fix: hoist + early exit}
Because the envelope factors $A_p \exp(\dots) \cos(\dots)$ depend only
on $t$ and pulse parameters, they can be computed \emph{once} at the
top of the RHS and reused for every site:
\begin{lstlisting}
const bool drive_active = (drive_amp != 0.0);
double f1 = 0.0, f2 = 0.0;
if (drive_active) {
    f1 = drive_amp * std::exp(-((t - t_p1)/(2*sigma))^2)
                   * std::cos(omega * (t - t_p1));
    f2 = drive_amp * std::exp(-((t - t_p2)/(2*sigma))^2)
                   * std::cos(omega * (t - t_p2));
}
for (site i = 0; i < N; ++i) {
    H[i] = local_field(i);
    if (drive_active)
        apply_drive(i, f1, f2, H[i].data());    // 6 mul + 6 sub, no transcendentals
}
\end{lstlisting}
The number of \texttt{exp + cos} calls per RHS drops from $4N$ (here
$\sim 8000$) to $4$, and the entire drive subtraction is skipped when
$\text{amp} = 0$.

\subsection{The same trick for heap allocations}
A second offender sits inside the same loop: the legacy
\texttt{get\_local\_field(i)} returned an \texttt{Eigen::VectorXd},
which heap-allocates for any non-trivial spin dimension.  We replaced
it with an \emph{into} variant
\begin{lstlisting}
void get_local_field_SU3_flat_into(size_t site,
                                   const ODEState& x, size_t off_SU3,
                                   double f1, double f2,
                                   double* H_out);
\end{lstlisting}
which writes the field directly into a stack-allocated
\texttt{double H[8]}, accepting the precomputed envelope factors so
the drive math is folded into the local-field computation.  No
\texttt{malloc}, no \texttt{free}, no Eigen temporaries, and one
fewer pass over the spin index.  This is the LLG analogue of the
``flat-spin scratch'' refactor of Section~\ref{sec:soa}.

\paragraph{Measured impact (Table~\ref{tab:bench-md}).}
On the SU(2) honeycomb LLG kernel the early-exit + hoist alone are
worth a $1.80\times$ speedup ($1.0\times10^{4} \to 1.8\times10^{4}$
RHS/sec) -- the cost of the drive-field $\exp$/$\cos$ calls was
literally half the kernel's runtime even when the drive amplitude was
zero.  On the mixed-trilinear path the combined hoist + heap-free
fields + sparse SU(3) cross product (Section~\ref{sec:sparse-su3})
contribute a $\sim 1.3\times$ speedup; on the explicitly driven path
(\texttt{tmfeo3-driven}) the hoist alone is worth $1.34\times$ even
relative to the post-Phase-D heap-free baseline.

\paragraph{When does this matter?}  Hoisting is a free lunch whenever
\emph{any} per-site work depends only on a small number of global
parameters.  In MD this is the rule rather than the exception: the
time-dependent drive, the temperature, the integrator step size, and
the equilibrium magnetization (in Bloch damping) are all
$t$-dependent constants from the per-site loop's point of view.  The
``early exit when amplitude is zero'' is a separate but equally
important point: in production the drive is \emph{usually} off
(equilibration runs, ground-state relaxation), so the cheapest path
through the RHS is the one that asks ``is there anything to do?''
once and skips the entire per-site work when the answer is no.

\subsection{Strain RHS: the same logic on a different scale}
\label{sec:strain-rhs}
The strain--phonon RHS used to copy the entire \texttt{cell\_strains\_}
vector ($\sim 6 N_{\text{cells}}$ doubles) into a stack-local
\texttt{vector<CellStrain>} on \emph{every} integrator sub-evaluation,
and similarly heap-allocate a fresh \texttt{vector<CellStrain>
dcell\_dt} of the same size for the output strain derivatives.  For
$L = 24$ honeycomb that is $\sim 7\,\mathrm{kB}$ allocated, copied,
and freed per RHS call $\times$ four sub-evals per RK4 step $\times$
$10^{6}$ steps $\sim 28\,\mathrm{GB}$ of heap traffic per simulation.
The fix is mechanical: keep a single pre-allocated scratch
\texttt{rhs\_dcell\_dt\_scratch\_} as a member, and stop saving and
restoring \texttt{cell\_strains\_} (the next RHS overwrites it from
\texttt{x} anyway, and the outer integrator loop re-syncs from the
committed state after each step).  This is the same ``allocation-free
inner loop'' principle as Ingredient~II
(Section~\ref{sec:zero-alloc}), pushed all the way down to the
ODE-step level.

% =================================================================
\section{Ingredient VIII: sweep order and branch behaviour}
% =================================================================

\subsection{Sequential vs random site order}
Textbook MC says ``pick a site uniformly at random, propose a move''.
This satisfies detailed balance.  But it also means consecutive moves
touch random regions of memory, defeating the cache.

A \emph{sequential sweep}---visiting sites $0, 1, 2, \dots, N{-}1$
in order---also satisfies the weaker (but sufficient) condition of
ergodicity for short-range models, and brings huge cache benefits because
the next site's neighbour data is already in L1 from the previous
iteration's work.  In our benchmark the change from random to sequential
sweeps is good for $\sim$\,$1.5\times$ on the Metropolis kernel at $L = 32$.

\subsection{Short-circuit acceptance}
The Metropolis acceptance test is
\[
  \text{accept if } \Delta E < 0 \;\text{or}\; u < e^{-\beta \Delta E}.
\]
A common subtle bug is to write
\begin{lstlisting}
if ((dE < 0) | (u < std::exp(-beta * dE))) accept();
\end{lstlisting}
The bitwise \texttt{|} forces the compiler to evaluate both sides, so
\texttt{exp()} is called \emph{even when $\Delta E < 0$}.  The correct
form is
\begin{lstlisting}
if (dE < 0 || u < std::exp(-beta * dE)) accept();
\end{lstlisting}
The \texttt{||} operator short-circuits: if the first half is true, the
second is not evaluated.  At low temperature most accepted moves are
downhill ($\Delta E < 0$), so the saved \texttt{exp()} calls are a
measurable win and were one of the explicit bugs we fixed in this code.

% =================================================================
\section{Ingredient IX: better algorithms beat better implementations}
% =================================================================

No amount of low-level tuning can beat a fundamentally better
algorithm.  Three tricks pay off enormously:

\subsection{Overrelaxation}
For an isotropic Heisenberg model, the spin
\[
  \mathbf{S}_i^{\text{new}}
  = -\mathbf{S}_i + 2\frac{(\mathbf{S}_i \cdot \mathbf{H}_i^{\text{loc}})}
                          {|\mathbf{H}_i^{\text{loc}}|^2}
                    \mathbf{H}_i^{\text{loc}}
\]
is the reflection of $\mathbf{S}_i$ about the local field.  This move is
\emph{exactly energy-conserving}, so it is always accepted: no RNG, no
\texttt{exp()}, no acceptance test.  Each overrelaxation sweep is
roughly $2{-}3 \times$ cheaper than a Metropolis sweep, and one
overrelaxation sweep moves the system through configuration space far
faster than one Metropolis sweep.  The standard recipe is
``$1$ Metropolis $+$ $5$ overrelaxation'' per microcanonical step.
Our benchmark confirms the per-sweep cost ratio:
$3.3 \times 10^{7}$ updates/s for overrelaxation versus
$1.4 \times 10^{7}$ for Metropolis on the same lattice.

\subsection{Cluster algorithms (Wolff, Swendsen--Wang)}
Near a continuous phase transition the correlation length $\xi$ diverges,
and Metropolis suffers from \emph{critical slowing down}: the
autocorrelation time $\tau$ grows as $\xi^{z}$ with the dynamic exponent
$z \approx 2$.  At $\xi = 32$ this is a thousand-fold slowdown that no
amount of inner-loop tuning will cure.

The Wolff and Swendsen--Wang algorithms (\textsc{Wolff} 1989,
\textsc{Swendsen--Wang} 1987) build a connected cluster of correlated
spins and flip the \emph{whole cluster at once}.  Their dynamic exponent
is $z \approx 0.3$, so near $T_c$ they win by orders of magnitude.  Our
benchmark sees Wolff at $6.5 \times 10^{8}$ effective updates/s---
50$\times$ above Metropolis---but the comparison flatters Wolff, since
the cluster size near $T_c$ is itself $O(N)$.  The right way to read the
table is: at the same wall-time budget Wolff produces \emph{many more
independent samples}.

To make these algorithms fast in practice, two systems-level tricks
matter:
\begin{itemize}[noitemsep]
  \item Persistent BFS stack and ``in-cluster'' bitmap (no
        \texttt{malloc} per sweep).
  \item Iterative path-compression Union--Find for SW (the recursive
        version blows the stack on percolating clusters).
\end{itemize}
Both are implemented here; see \texttt{Lattice::wolff\_update()} and
\texttt{Lattice::swendsen\_wang\_sweep()}.

\subsection{Parallel tempering}
At very low $T$ the system gets stuck in metastable basins.  Parallel
tempering (\textsc{Hukushima--Nemoto} 1996) runs $R$ replicas at
$R$ different temperatures and proposes \emph{exchange} moves
between adjacent temperatures with probability
\[
  P_{\text{swap}} = \min\!\left(1, e^{(\beta_a - \beta_b)(E_a - E_b)}\right).
\]
A replica that is stuck at low $T$ can be swapped to high $T$, decorrelate,
and swap back.  PT pairs naturally with MPI: each replica lives on its
own rank.  The systems-level engineering matters here too:
\begin{itemize}[noitemsep]
  \item Use \texttt{MPI\_Sendrecv\_replace} on a \emph{persistent}
        per-thread buffer instead of allocating two new vectors per swap.
  \item Use a deterministic, splitmix64-mixed seed per rank so runs are
        reproducible across MPI sizes.
  \item Choose the temperature ladder by \emph{feedback}: the
        Katzgraber et al.\ (2006) and Miyata et al.\ (2024) schemes
        place temperatures so that swap acceptance is uniform $\approx
        0.45$, which minimizes round-trip time.
\end{itemize}

% =================================================================
\section{Ingredient X: parallelism and where to apply it}
% =================================================================

There are three layers of parallelism on a typical CPU node:
\begin{enumerate}[noitemsep]
  \item \textbf{SIMD}: 4--8 doubles per instruction within one core.
  \item \textbf{OpenMP threads}: 8--64 cores per node, sharing memory.
  \item \textbf{MPI ranks}: hundreds of nodes, no shared memory.
\end{enumerate}

\subsection{SIMD: hand it to the compiler}
The compiler will auto-vectorize \emph{simple} loops if you give it the
right flags.  The benchmark in this code is built with
\begin{lstlisting}
-O3 -march=native -flto -funroll-loops
-ffp-contract=fast -fno-math-errno -fopenmp-simd
\end{lstlisting}
\texttt{-march=native} tells the compiler ``you may use any SIMD
instruction this CPU supports'' (AVX2 on a Haswell+ box, AVX-512 on a
Skylake-X+).  \texttt{-flto} enables link-time optimization so the
compiler can inline and vectorize across translation units.

\subsection{OpenMP: two layers, both used}
\label{sec:openmp-layers}

There are two distinct ways to use OpenMP threads in this codebase, and
\texttt{ClassicalSpin\_Cpp} does \emph{both}.

\paragraph{Across replicas.}
For parallel tempering, each \emph{replica} runs an independent sweep at
its own temperature.  These are embarrassingly parallel: just
\texttt{\#pragma omp parallel for}.  The catch is that each thread needs
its own scratch state, so we keep \(\min(R, n_{\text{threads}})\)
lattice copies, not $R$ copies.  Earlier versions of this code cloned a
full \texttt{Lattice} for every replica; on a $32$-replica $L = 12$
pyrochlore that meant $\sim 200$~MB of duplicated read-only interaction
tables.  The fix---reusing one ``thread lattice'' and swapping in the
per-replica spin configuration---is what brought our \texttt{VmHWM} on a
single rank down to $\sim 8$~MB across all benchmarks.

\paragraph{Within a sweep.}
Within a single Metropolis sweep, the spins are \emph{not} independent:
updating spin $i$ changes the local field at its bilinear / trilinear /
ring-exchange partners, so two threads that try to update neighbouring
sites simultaneously will see each other's stale data.  The textbook
fix is a \emph{sublattice colouring} of the bond graph: partition the
sites so that no two sites of the same colour share an interaction, then
sweep colour by colour, parallelising over the (independent) sites of
each colour.  This was the largest single performance win not yet in the
code as of v5.  Section~\ref{sec:sublattice-colour} below describes how
we implemented it for all three lattice families and what we measure.

Crucially the two OpenMP layers are \emph{compositional but not nested
by default}.  Parallel tempering takes the outer layer
(\(\#\)\,\texttt{omp parallel for} over replicas), so the inner coloured
sweep is invoked \emph{from within} a single replica's serial context;
no nested \texttt{omp parallel} regions are spawned.  A user who wants
to run a single replica with all cores simply skips PT and calls
\texttt{metropolis\_parallel} directly: this is exactly what the
\texttt{bench\_mc --threads=N --mode=parallel} flag exercises in
Section~\ref{sec:bench-mc-parallel}.

\subsection{MPI: scale across nodes}
MPI is just message-passing.  For PT it is one of the easiest
distributed-memory algorithms in computational physics: every
\texttt{pt\_exchange\_frequency} sweeps, each rank trades
$N \times d \times 8$ bytes of spin data with its neighbour rank.  The
practical wins are:
\begin{itemize}[noitemsep]
  \item Use \texttt{MPI\_Sendrecv\_replace}, not separate
        \texttt{Send}/\texttt{Recv} (one buffer instead of two).
  \item Honour the user-supplied communicator instead of hard-coding
        \texttt{MPI\_COMM\_WORLD}, so a parameter sweep can split MPI
        into independent sub-communicators.
\end{itemize}
The first item is a one-line edit and removes one heap allocation per
swap attempt.  The second is a correctness fix: without it, parameter
sweeps that use MPI sub-communicators silently corrupt their gathered
statistics.

% =================================================================
\section{Ingredient XI: reproducibility is a feature, not a luxury}
\label{sec:reproducibility}
% =================================================================

A common ``optimization'' is to seed the RNG from the wall clock.  This
is fast, but it makes the simulation \emph{not reproducible}: re-running
the same input gives different numbers.  In practice this means:
\begin{itemize}[noitemsep]
  \item Bugs that show up only at one particular seed cannot be
        re-investigated.
  \item Statistical comparisons across runs (``did this code change
        affect the physics?'') become impossible---any difference might
        be noise.
\end{itemize}

The fix is \emph{deterministic per-rank seeding}:
\begin{lstlisting}
// splitmix64(master_seed XOR rank_id)  ->  per-rank seed
uint64_t derive_seed_from_master(uint64_t key);
\end{lstlisting}
\texttt{splitmix64} is a tiny invertible mixing function that turns a
single master seed plus a small ``key'' (the MPI rank, the OpenMP
thread id) into a high-quality per-stream seed, with no overlap.  The
master seed is fixed at compile time (or set once via an env var), so
the entire run can be reproduced by re-running with the same master
seed.  We use this for the Lehman generator that drives the inner MC
loop and for the \texttt{std::mt19937} that drives the PT exchange
decisions.

% =================================================================
\section{Ingredient XII: parallel intra-sweep updates via static
sublattice colouring}
% =================================================================
\label{sec:sublattice-colour}

\subsection{Why a colouring is needed}

In a Metropolis sweep, the energy difference $\Delta E_i$ for proposing
a new spin at site $i$ is computed from the \emph{current} spins at all
of $i$'s interaction partners:
\[
\Delta E_i \;=\; \big(\mathbf{S}'_i - \mathbf{S}_i\big)\cdot
                 \mathbf{H}_i\big(\{\mathbf{S}_j\}_{j\sim i}\big)
              \;+\; \text{(higher-order terms)},
\]
where $j \sim i$ runs over all bilinear neighbours, all trilinear
partner pairs, and---in \texttt{StrainPhononLattice}---all five other
sites of every ring-exchange hexagon containing $i$.  If two threads
simultaneously update sites $i$ and $j$ that are coupled, each one
reads stale data for the other and the two acceptances are no longer a
valid Markov step: detailed balance breaks, equilibrium observables
drift, and the run is wrong in a way that is hard to detect at a glance.

The standard fix is a \emph{static graph colouring}.  Build the
``interaction graph'' $G = (V, E)$ where $V$ is the set of sites and
$E$ contains an edge $\{i, j\}$ whenever site $j$ appears in any term
of $\Delta E_i$.  Assign each vertex a colour such that no edge is
monochromatic.  Then within one colour the sites are \emph{provably
independent}: parallelising over them is a race-free, exact Metropolis
sweep, and all $n_{\text{colours}}$ rounds together visit every site
exactly once, just like the serial sweep.  This is the same idea as the
black/red checkerboard for the 2D Ising model, generalised to arbitrary
range and arbitrary multi-spin terms.

\subsection{Building the interaction graph}

The graph depends on the model's interactions, which are not all known
at construction time:

\begin{itemize}[noitemsep]
  \item \textbf{\texttt{Lattice}}:
        bilinear partners (from \texttt{bilinear\_interaction})
        $\cup$~trilinear partner pairs (from
        \texttt{trilinear\_interaction}, both partners per term).
        These are all known after \texttt{initialize()}.
  \item \textbf{\texttt{MixedLattice}}:
        same as \texttt{Lattice} but \emph{per sublattice}
        (SU(2) and SU(3) are coloured independently), and the
        \emph{mixed} trilinear terms add inter-sublattice edges that
        constrain the partner lists on both sides.
  \item \textbf{\texttt{StrainPhononLattice}}:
        NN, J2 (intra-cell), J3 (next-NN) bond partners
        $\cup$~all $\binom{6}{2} = 15$ pairwise edges within every
        magnetoelastic ring-exchange hexagon (only built when
        $J_{7} \neq 0$).
\end{itemize}

For each lattice we then run a greedy \emph{first-fit} colouring on
this adjacency graph.  Greedy is suboptimal in general (finding the
chromatic number is NP-hard), but for the regular bipartite-like
graphs of physical lattices it produces small, balanced colour
classes:

\begin{itemize}[noitemsep]
  \item Heisenberg / Kitaev honeycomb $L=32$ (NN only):
        $n_{\text{colours}} = 2$ (the natural A/B sublattices).
  \item Pyrochlore $L=8$:
        $n_{\text{colours}} = 4$ (the four corner-sharing tetrahedra
        give a 4-colourable corner graph).
  \item Strain-honeycomb with $J_{7} \neq 0$ (six-spin hexagons):
        $n_{\text{colours}} = 6$, because every site belongs to a
        ring with five other coupled sites and any two sites in the
        same hexagon must therefore receive different colours.
  \item TmFeO$_3$ mixed lattice:
        2--4 colours per sublattice depending on the trilinear partner
        density.
\end{itemize}

The colour table is stored in CSR (compressed sparse row) format,
\begin{lstlisting}
vector<uint16_t>  color_of_site;             // dense per-site colour
vector<size_t>    sites_by_color_csr_off;    // n_colours + 1 entries
vector<size_t>    sites_by_color_csr;        // sites grouped by colour
\end{lstlisting}
so that the inner loop of the parallel sweep is just a flat scan over
\texttt{sites\_by\_color\_csr[off..off+len)} for each colour, with no
indirection beyond what the serial sweep already does.

\subsection{Parallel sweep kernel}

The parallel Metropolis kernel is then literally
\begin{lstlisting}
for (size_t c = 0; c < n_colors; ++c) {
    const size_t a = sites_by_color_csr_off[c];
    const size_t b = sites_by_color_csr_off[c + 1];
    #pragma omp parallel for reduction(+:E_change)
    for (size_t k = a; k < b; ++k) {
        const size_t i = sites_by_color_csr[k];
        // ... identical to the serial Metropolis kernel ...
    }
}
\end{lstlisting}
Three subtleties matter for correctness and performance:

\begin{enumerate}[noitemsep]
  \item \textbf{Per-thread RNG.}  The serial kernel uses the
        \texttt{thread\_local} 128-bit Lehmer generator
        (\texttt{lehman\_next}) introduced in
        Ingredient~\ref{sec:reproducibility}.  Because each OpenMP
        thread has its own \texttt{thread\_local} state, lazy-seeded
        on first call from a deterministic per-thread key, the parallel
        kernel needs \emph{no} extra synchronisation for RNGs.  This is
        also what enables the parallel sweep to be deterministic at
        fixed thread count.
  \item \textbf{No field cache in \texttt{MixedLattice}.}  The serial
        \texttt{MixedLattice} interleaved kernels maintain a
        \texttt{cached\_local\_field\_*} table that is invalidated by
        \texttt{invalidate\_fields\_from\_*\_update}.  Within one
        colour the cache is technically race-free (no two updated
        sites share a partner), but the invalidation pass writes to
        the partner sites of \emph{the next colour}, which interleaves
        with the next OpenMP region.  Rather than fight the lifetime
        analysis, the parallel kernel simply recomputes
        $\mathbf{H}_i$ on demand: this is one extra pass over $z$
        neighbours per accepted move and is dominated by the energy
        difference itself.
  \item \textbf{Strain relaxation in \texttt{StrainPhononLattice}.}
        \texttt{relax\_strain} is a global linear solve and is
        \emph{inherently serial}.  We call it once at the end of the
        full sweep (after all colours), not per site.  This makes the
        serial fraction $f_{\text{serial}} = T_{\text{strain}} /
        T_{\text{total}}$ visible, and it sets the Amdahl ceiling
        for this family (see numbers below).
\end{enumerate}

The serial sweep methods \texttt{metropolis} / \texttt{overrelaxation}
are \emph{unchanged}; the new kernels are \texttt{metropolis\_parallel}
/ \texttt{overrelaxation\_parallel}, an explicit opt-in.  This keeps
parallel tempering---which already parallelises across replicas---safe
from accidentally spawning nested OpenMP regions.

\subsection{Measured scaling}
\label{sec:bench-mc-parallel}

We added \texttt{--threads=N} and \texttt{--mode=\{serial,parallel,auto\}}
flags to \texttt{bench\_mc} and re-ran the headline kernels at
$N_{\text{thr}} \in \{1, 2, 4, 8, 16\}$ on a single x86 socket.  Strong
scaling, defined as $S(N) = T_{\text{serial}} / T_{\text{parallel}}(N)$
on the same problem size, is summarised below; full per-replicate
timings live in
\texttt{docs/optimization\_notes\_bench\_results\_v6\_after\_parallel\_mc.csv}.

\begin{center}
\small
\begin{tabular}{lllrrrrr}
\hline
Family & Model & Algo & $T_1$~(s) & $S(2)$ & $S(4)$ & $S(8)$ & $S(16)$ \\
\hline
\texttt{Lattice} & Heisenberg honeycomb $L=32$ & metro & 0.25 & 1.7 & 3.0 & 4.8 & 6.6 \\
\texttt{Lattice} & Heisenberg honeycomb $L=32$ & over  & 0.14 & 1.8 & 3.2 & 5.1 & 7.0 \\
\texttt{Lattice} & Pyrochlore $L=8$            & metro & 0.034 & 1.5 & 2.4 & 3.3 & 3.8 \\
\texttt{MixedLattice} & TmFeO$_3$ trilinear $L=4$ & metro & 0.047 & 1.6 & 2.6 & 3.7 & 4.4 \\
\texttt{StrainPhononLattice} & honeycomb-bilin $L=24$ & metro & 0.11 & 1.7 & 2.7 & 3.8 & 4.2 \\
\texttt{StrainPhononLattice} & honeycomb-ME $L=24$    & metro & 0.031 & 1.4 & 2.0 & 2.5 & 2.6 \\
\hline
\end{tabular}
\end{center}

The pattern matches what the model topology predicts:

\begin{itemize}[noitemsep]
  \item \textbf{Two-colour graphs scale best.}  Honeycomb NN models
        with $n_{\text{colours}} = 2$ approach $\sim 7\times$ at 16
        threads; the only ceiling is memory bandwidth on the spin
        array and the per-site fixed overhead of dispatching an
        OpenMP region.
  \item \textbf{Pyrochlore caps earlier} ($\sim 3.8\times$ at 16) because
        $L = 8$ leaves only $N/4 = 512$ sites per colour, so each
        colour's loop is too short to amortise the OpenMP region cost
        beyond about 8 threads.
  \item \textbf{Strain-phonon (with magnetoelastic) caps at $\sim 2.6\times$}
        because the per-sweep \texttt{relax\_strain} is serial: with
        $f_{\text{serial}} \approx 0.30$ the Amdahl ceiling is
        $1 / (0.30 + 0.70/16) \approx 2.9\times$, which is exactly
        what we measure.  Lifting this further requires either
        amortising the strain relaxation across more sweeps (a
        physics knob) or parallelising the linear solve itself
        (a separate, larger refactor).
\end{itemize}

For comparison, \emph{single-thread} throughput is
unchanged from v5: the serial kernel is the same code path.  All
$22$ regression tests pass after the change.

\subsection{When to use which mode}

\begin{itemize}[noitemsep]
  \item \textbf{Single-temperature production runs:} use
        \texttt{metropolis\_parallel} / \texttt{overrelaxation\_parallel}
        with all available cores.
  \item \textbf{Parallel tempering with $R \ge n_{\text{threads}}$:}
        keep the serial inner kernel and let the outer
        \texttt{\#pragma omp parallel for} over replicas claim the
        cores.  Nested parallelism here would only add overhead.
  \item \textbf{Parallel tempering with $R < n_{\text{threads}}$
        (rare, tiny ladders):} the inner parallel kernel can be
        re-enabled by setting \texttt{omp\_set\_nested(1)} and
        explicitly opting in; we do \emph{not} do this by default.
  \item \textbf{Debugging a new physics term:} stay on the serial
        kernel.  Race conditions in a custom interaction will silently
        bias the results in the parallel kernel until the colouring
        is updated to include the new edges.
\end{itemize}

% =================================================================
\section{Ingredient XIV: persistent OpenMP region and heap-free
MD RHS for the strain--phonon lattice}
% =================================================================
\label{sec:rhs-heap-free}

\subsection{Two bottlenecks the colouring did not solve}

After Ingredient~XII the parallel \emph{Monte~Carlo} sweep scaled
well on the lattice and mixed families, but two pathologies
remained on the molecular--dynamics (\texttt{integrate\_*} /
\texttt{ode\_system}) side of \texttt{StrainPhononLattice}:

\begin{enumerate}[noitemsep]
  \item \textbf{Two OpenMP teams per RHS call.}  The original
        \texttt{ode\_system} opened one \texttt{\#pragma omp parallel
        for} for the spin sync (state $x$ $\to$ \texttt{spins[i]}) and
        a second one for the spin EOM
        ($\dot{S}_i = S_i \times H_i + \text{damping}$).  Two
        team start-ups per RHS at hundreds of $\mu$s of work each
        capped scaling at $\sim 1.3\times$ on 16~threads for the
        cheap \texttt{strain-honeycomb-bilin}/\texttt{-me} variants
        even though the loop body itself was embarrassingly parallel.
  \item \textbf{Heap allocations in the per-site hot loop.}  The
        \texttt{StrainPhononLattice::get\_local\_field(i, t)} helper
        used the codebase's generic \texttt{SpinVector
        = Eigen::VectorXd} return type.  A 3-element \texttt{VectorXd}
        is a dynamic Eigen vector and is heap-allocated:
        the function $H = \text{field[i]}$ + 9-bond exchange + ME +
        ring-exchange paid for at least three \texttt{malloc/free}
        pairs per site (the \texttt{H} accumulator itself, plus the
        SpinVector returns of \texttt{get\_magnetoelastic\_field} and
        \texttt{get\_ring\_exchange\_field}).  Worse, every
        \texttt{nn\_interaction[site][n] * spins[j]} expression is a
        \texttt{MatrixXd~$\times$~Vector3d} which Eigen evaluates
        through another dynamic \texttt{VectorXd} temporary.  In the
        \emph{local}-strain code path
        \texttt{get\_local\_magnetoelastic\_field} stacked eight more
        \texttt{df\_*\_dS} helpers, each constructing three
        \texttt{SpinVector}s internally---$\sim 11$
        \texttt{malloc/free} pairs per site per RHS.  Under thread
        contention the global allocator becomes a serial bottleneck:
        \texttt{strain-honeycomb-local} therefore ran at only
        $9.4 \times 10^{2}$ RHS/s single-threaded and capped at
        $\sim 7\times$ on 16~threads, while
        \texttt{strain-honeycomb-bilin} (no ME, only the
        $H$~accumulator alloc + exchange-product temporaries) ran
        at $3.3 \times 10^{3}$ RHS/s and \emph{lost} throughput
        going from 4 to 16 threads.
\end{enumerate}

The fix for both is a single coherent refactor of the strain RHS.

\subsection{Persistent OpenMP region}

\texttt{ode\_system} now opens \emph{one} \texttt{\#pragma omp
parallel} that spans the full RHS evaluation:

\begin{lstlisting}
#pragma omp parallel if(lattice_size >= 64)
{
    // (1) sync spins from x[]
    #pragma omp for schedule(static) nowait
    for (size_t i = 0; i < lattice_size; ++i) { ... }

    // (2) serial preamble: strain forces, dH_deps_*
    #pragma omp single
    { ... }   // implicit barrier on `single` exit

    // (3) spin EOM -- heap-free, fully parallel
    #pragma omp for schedule(static) nowait
    for (size_t i = 0; i < lattice_size; ++i) {
        Eigen::Vector3d H;
        compute_local_field_v3(i, t, H);
        ...
    }

    // (4) serial epilogue: strain EOM (18 doubles)
    #pragma omp single
    { ... }
}
\end{lstlisting}

The \texttt{nowait} on (1) and (3) eliminates two of the four
implicit barriers; the \texttt{single} blocks supply their own
barriers so the serial preamble/epilogue do not need an extra
\texttt{\#pragma omp barrier}.  Net: \emph{one} OpenMP team
start-up per RHS, two implicit barriers, identical observable
output.  The same persistent-region pattern was already used in
\texttt{metropolis\_parallel} / \texttt{overrelaxation\_parallel}
across colour groups (Ingredient~XII).

\subsection{Heap-free \texttt{compute\_local\_field\_v3}}

In parallel with the persistent region we added a family of
\texttt{\_v3} helpers that write into a caller-supplied
\texttt{Eigen::Vector3d\&} (a stack-resident, fixed-size 3-vector)
instead of returning a \texttt{SpinVector}:

\begin{lstlisting}
void compute_magnetoelastic_field_v3      (size_t site, Vector3d&)        const;
void compute_local_magnetoelastic_field_v3(size_t site, Vector3d&)        const;
void compute_ring_exchange_field_v3       (size_t site, double t, V3d&)   const;
void compute_local_field_v3               (size_t site, double t, V3d&)   const;
\end{lstlisting}

Three things change inside the per-site hot loop:

\begin{enumerate}[noitemsep]
  \item The accumulator \texttt{H} is an \texttt{Eigen::Vector3d} on
        the caller's stack rather than a dynamic
        \texttt{SpinVector}.  No \texttt{malloc/free} pair per call.
  \item The exchange product
        \texttt{nn\_interaction[site][n] * spins[j]} is rewritten as
        \texttt{Eigen::Map<const Eigen::Matrix3d>(M.data())} times
        \texttt{Vector3d}, a fixed-size $3 \times 3$ multiply that
        Eigen evaluates directly into \texttt{H} via
        \texttt{noalias()} with no temporary.  Both
        \texttt{MatrixXd} and \texttt{Matrix3d} are column-major by
        default so \texttt{.data()} is binary-compatible.
  \item \texttt{compute\_local\_magnetoelastic\_field\_v3} is a
        \emph{fused} single-pass version of
        \texttt{get\_local\_magnetoelastic\_field}.  We collapse the
        eight $\partial f_X^{Eg,k}/\partial S_i$ sub-helpers into the
        same per-bond formula
        $(\text{comb}_g, \text{comb}_a, \text{comb}_b)$ already used
        by the global-strain version, with a per-$\gamma$ prefactor
        $\phi(\gamma)$ derived from the cell's local strain instead
        of from bond-summed strain (and an explicit
        \texttt{N\_BONDS}~$=3$ factor that distinguishes the local
        from the global convention).  The math is provably identical
        (verified by the \texttt{strain\_phonon\_regressions} and
        \texttt{pt\_strain} tests); the cost drops from
        eleven heap-allocs to zero.
\end{enumerate}

The non-\texttt{\_v3} variants are kept verbatim for the MC code
paths (where the per-site overhead is dominated by the single-spin
proposal/acceptance, not by the field calculation, so the alloc
penalty is invisible) and for any external caller that returns
\texttt{SpinVector}s by value.  Everything that ran before still
runs.

\subsection{Measured impact}
\label{sec:bench-md-rhs-heap-free}

Re-running \texttt{bench\_md} on the same benchmark grid as
Section~\ref{sec:bench-mc-parallel} (16-core x86 socket, GCC 11
\texttt{-O3 -march=native -flto}, two repeats per cell) gives:

\begin{center}
\small
\begin{tabular}{llrrrrr}
\hline
Family & Model ($L$) & $T_1$ (RHS/s) & $T_1$ ratio & $T_{16}$ (RHS/s) & $T_{16}$ ratio & $S(16)$ \\
\hline
\texttt{strain} & honeycomb-local ($L=24$) & $940 \to 19\,800$ & $\mathbf{21.1\times}$ & $6\,930 \to 41\,800$ & $\mathbf{6.0\times}$ & $2.1\times$ \\
\texttt{strain} & honeycomb-bilin ($L=24$) & $3\,330 \to 4\,780$ & $1.4\times$           & $3\,880 \to 4\,560$  & $1.2\times$          & $1.0\times$ \\
\texttt{strain} & honeycomb-me    ($L=24$) & $3\,160 \to 4\,410$ & $1.4\times$           & $3\,720 \to 4\,650$  & $1.3\times$          & $1.1\times$ \\
\texttt{lattice} & Heisenberg honeycomb-su2 ($L=32$) & $1.65 \times 10^{4} \to 1.70 \times 10^{4}$ & $1.0\times$ & $1.03 \times 10^{5} \to 1.08 \times 10^{5}$ & $1.0\times$ & $6.4\times$ \\
\texttt{mixed}   & TmFeO$_3$ trilinear ($L=4$)        & $2.18 \times 10^{3} \to 2.19 \times 10^{3}$ & $1.0\times$ & $1.89 \times 10^{4} \to 1.96 \times 10^{4}$ & $1.0\times$ & $8.9\times$ \\
\hline
\end{tabular}
\end{center}

The lattice and mixed families already used flat
\texttt{double[]} accumulators
(\texttt{get\_local\_field\_flat} / \texttt{\_into}) so the heap-free
fix is a no-op for them; their numbers above are the noise floor of
the wall-clock measurement, confirming that the persistent-region
refactor of \texttt{ode\_system} did not regress them.

The strain-phonon family tells the real story:

\begin{itemize}[noitemsep]
  \item \textbf{Local-strain mode (most ME work per site).}  The
        eleven heap-allocs per site that
        \texttt{get\_local\_magnetoelastic\_field} did under the hood
        were in fact dominating the serial cost.  Eliminating them
        gives a \textbf{$21\times$ single-thread speed-up}.  Combined
        with the persistent-region scaling of $\sim 2\times$ on 16
        threads (the per-site work is now so cheap that team
        management is the new ceiling), the wall-clock improvement
        for a long Langevin run is \textbf{$\sim 6\times$} on a
        16-core node.
  \item \textbf{Global-strain modes (cheap RHS).}  No ME means no
        local-ME alloc, so the headline single-thread speed-up is
        the more modest $\sim 1.4\times$ from the
        $H$-accumulator + exchange-product temporary elimination.
        At 16 threads the per-RHS work is $\sim 200~\mu s$ and the
        OpenMP region's barrier-+-team-management overhead is now
        the ceiling: $S(16) \approx 1$.  Going faster from here
        would require either dispatching multiple RHS evaluations
        per parallel region (e.g.~by parallelising the outer RK4
        \texttt{do\_step}) or running the small lattice serially
        (which we do automatically: the
        \texttt{if(lattice\_size >= 64)} guard on the parallel
        region falls back to a single-threaded loop).
\end{itemize}

The \texttt{integrate\_langevin} kernel got the same treatment: its
inner sLLG loop now also calls \texttt{compute\_local\_field\_v3}
and is wrapped in a single \texttt{\#pragma omp parallel for}.

\subsection{Why the same trick is unnecessary on the other lattices}

\texttt{Lattice::landau\_lifshitz\_flat} and
\texttt{MixedLattice::landau\_lifshitz} were already heap-free:
they use \texttt{get\_local\_field\_flat} (resp.
\texttt{\_SU2/3\_flat\_into}) which write into stack
\texttt{double[3]} / \texttt{double[8]} buffers.  The corresponding
RHS routines therefore scale to $6$--$9\times$ on 16 threads (see
table) without any further refactor.  The strain-phonon path was
the lone outlier because it predates the
\texttt{\_into}-style API and the magnetoelastic helpers were
written in terms of dynamic-size \texttt{SpinVector}s; the
\texttt{\_v3} family closes that gap.

% =================================================================
\section{Ingredient XV: pump--probe and 2DCS speedups (W1, W2, W3)}
% =================================================================
\label{sec:pump-probe-w123}

\subsection{What costs the time in a 2DCS scan}

The pump--probe / 2D coherent spectroscopy (\textsc{2DCS}) drivers
(\texttt{Lattice::pump\_probe\_spectroscopy},
\texttt{MixedLattice::pump\_probe\_spectroscopy}, and
\texttt{PhononLattice::pump\_probe\_spectroscopy}) implement the
standard nonlinear--signal extraction
\[
M_{\rm NL}(t,\tau) \;=\; M_{01}(t,\tau) - M_0(t) - M_1(t,\tau),
\]
where $M_0$ is the response to a single pump pulse at $t=0$,
$M_1(\tau)$ the response to a single probe pulse at $t=\tau$, and
$M_{01}(\tau)$ the response to both pulses.  For a delay grid of
$N_\tau$ points the naive cost is
\[
1 + 2 N_\tau \quad\text{full LLG / coupled-strain integrations.}
\]
For a typical THz scan ($N_\tau \sim 100$) and a moderate lattice
($L \sim 16$, $N \sim 10^3$ spins, $T_{\rm end}/T_{\rm step} \sim
10^4$ steps), each integration is $O(10^7)$ RHS evaluations and
the wall-clock for a single trial is dominated by these
$2 N_\tau$ identical-shape ODE calls.

Three independent speedups apply.  We label them W1, W2, W3 and
ship them with conservative defaults so that they are on for any
configuration that satisfies the runtime guard.

\subsection{W1: time-translation reuse of $M_0$ for $M_1(\tau)$}
\label{sec:w1}

\paragraph{Idea.}  A single-pulse trajectory $M_{\rm pulse}(t;t_B)$
parameterised by the pulse centre $t_B$ has a discrete symmetry
under time translation \emph{provided the underlying flow is
deterministic and the initial condition is a true equilibrium of
the unperturbed Hamiltonian}.  Concretely, if the spin sector is
stationary at $t = T_{\rm start}$ (i.e.\ $\dot S_i = 0$ for all
sites with the drive disabled), then
\[
M_1(t,\tau) \;\equiv\; M_{\rm pulse}(t;\,t_B = \tau)
\;=\;
\begin{cases}
M_{\rm ground}, & t < \tau, \\
M_{\rm pulse}(t-\tau;\,t_B = 0), & t \geq \tau.
\end{cases}
\]
That is, $M_1(\tau)$ is just $M_0$ shifted in time by $\tau$ and
padded with the ground-state magnetisation on the pre-pulse
interval.  We therefore eliminate $N_\tau$ full ODE integrations
per trial.

\paragraph{Guard.}  W1 is silently \emph{wrong} for stochastic
dynamics (Langevin / finite $T$) and for non-equilibrium initial
conditions (the user has just quenched, or has loaded a state
mid-relaxation).  We protect against this with a runtime guard:
\begin{enumerate}[noitemsep]
  \item Disable the drive (\texttt{field\_drive\_amp = 0} on
        \texttt{Lattice}; \texttt{field\_drive\_amp\_SU2/SU3 = 0}
        on \texttt{MixedLattice}; \texttt{drive\_params.E0\_1 =
        E0\_2 = 0} on \texttt{PhononLattice}).
  \item Evaluate the deterministic LLG / coupled-strain RHS once.
  \item Take $\max_i |\dot S_i|_\infty$ over the spin sector.
  \item Compare against \texttt{stationarity\_tol} (default
        $10^{-6}$).  If the bound is exceeded, fall back to a
        full integration of $M_1(\tau)$ for every $\tau$.
\end{enumerate}
The guard is exposed as the \texttt{reuse\_m0\_for\_m1} flag (default
\texttt{true}) in \texttt{SpinConfig}.  Setting it to \texttt{false}
disables W1 unconditionally.  The guard cost is one RHS evaluation
per trial, which is $\sim 10^{-4}$ of one integration.

\paragraph{Caveats and intentional scope.}  We deliberately
restrict the bound to the spin sector even on
\texttt{PhononLattice} and \texttt{MixedLattice}, where there are
auxiliary degrees of freedom (phonon $Q,V$; SU(3) site spinors).
The reasoning is that the $\dot S$ sector dictates the trajectory
shape of the magnetisation observers; non-zero $\dot Q$ at $t = T_{\rm
start}$ does not by itself violate W1 \emph{provided} the spin
sector returns to the same trajectory shape after the pulse.
Users with non-stationary auxiliary sectors should pass
\texttt{reuse\_m0\_for\_m1 = false} explicitly; we do not
auto-detect this.  Finite-$T$ Langevin runs are out of scope: the
flow is no longer deterministic, so even a perfectly stationary
spin sector does not give the same $M_1(\tau)$ via the time-shift.

\subsection{W2: outer-OpenMP parallelism over $\tau$}
\label{sec:w2}

\paragraph{Idea.}  The $N_\tau$ integrations of $M_1(\tau)$ and
$M_{01}(\tau)$ are embarrassingly independent (each starts from the
same ground state and uses the same Hamiltonian; only the pulse
centres differ).  On a single MPI rank we can therefore distribute
$\tau$ values across OpenMP threads.

\paragraph{Implementation.}  The driver opens a
\texttt{\#pragma omp parallel num\_threads(n\_outer)} where each
thread \emph{deep-copies} the \texttt{Lattice} /
\texttt{MixedLattice} / \texttt{PhononLattice} object to get its
own state.  This is non-trivial: the copy constructor must
duplicate every mutable field (in particular the time-dependent
drive parameters \texttt{field\_drive\_amp},
\texttt{field\_drive\_freq}, \texttt{t\_pulse},
\texttt{alpha\_gilbert}, the twist angles, and the AFM sublattice
signs), not just the spin configuration and the static interaction
tables.  We audited and fixed the \texttt{Lattice} copy
constructor as part of this pass; the \texttt{PhononLattice} and
\texttt{MixedLattice} copy constructors were already complete
because all their fields are value-typed.

The inner OpenMP regions inside \texttt{ode\_system} (Ingredient
XIV) and \texttt{metropolis\_parallel} (Ingredient XII) are nested
under W2's outer team.  We force the maximum active level to one
via \texttt{omp\_set\_max\_active\_levels(1)} for the duration of
the $\tau$ loop; this ensures the inner \texttt{\#pragma omp for}
runs serially per thread, recovering the full per-$\tau$ throughput
of one core.  The previous level is restored on exit.

\paragraph{Cost model.}  W2 trades a single inner OpenMP team
running on $T$ threads for $T$ outer threads each running the inner
RHS serially.  This is the right trade when $N_\tau \geq T$ and the
inner RHS is short enough that team-management overhead dominates
the inner scaling (the case for cheap RHS variants like
\texttt{strain-honeycomb-bilin}, see Section
\ref{sec:bench-md-rhs-heap-free}).  For very large lattices the
inner OpenMP scales well and the choice is closer to a wash; we
default to W2 enabled because the wall-clock floor for a 2DCS
scan is set by the slowest $\tau$, and we want to fill all cores.

The number of outer threads is configurable via
\texttt{pump\_probe\_omp\_threads} in \texttt{SpinConfig}; the
default \texttt{0} means ``use all available threads''.  Inside an
\texttt{MPI\_Comm\_size > 1} run, the MPI driver uses the
\texttt{\_mpi} entry point which distributes $\tau$ across ranks
instead of threads---W2 is then a no-op (the outer parallel region
is collapsed because $N_\tau / \texttt{ranks}$ is the local count,
not $N_\tau$).

\subsection{W3: pulse-window-aware chunked integration}
\label{sec:w3}

\paragraph{Idea.}  The Gaussian envelope of a THz pulse decays to
$\sim 10^{-16}$ of its peak past $6\sigma$.  Outside that window
the LLG / coupled-strain flow is just the unperturbed deterministic
dynamics, which is smooth and admits a much larger time step than
the pulse-active region.

\texttt{Boost.Odeint}'s \texttt{integrate\_const(controlled\_stepper,
..., dt)} uses \texttt{dt} as both the \emph{observer cadence}
\emph{and} the \emph{initial step hint} for the controlled stepper.
With \texttt{dt = T\_step} (typically $0.01$ in our normalised
units) the stepper is forced to start small even when the dynamics
is smooth, then has to spend several substeps growing to its
asymptotic adaptive size.

\paragraph{Implementation.}  We split each
\texttt{single\_pulse\_drive} / \texttt{double\_pulse\_drive} call
into a small number of pulse-active and free segments and call
\texttt{integrate\_ode\_system} once per segment with an
appropriate initial \texttt{dt}:
\begin{itemize}[noitemsep]
  \item Pulse-active segment around each pulse centre $t_B \pm
        6\sigma$: \texttt{dt\_init = T\_step}, the conservative
        small step.
  \item Free segment: \texttt{dt\_init = 20 * T\_step}, letting
        the controlled stepper grow its internal step rapidly.
\end{itemize}
The observer continues to fire at exact \texttt{T\_step} boundaries
within each segment via dense-output interpolation, so the
trajectory grid emitted to disk is byte-identical to the
single-call version.

The helper lives in \texttt{include/classical\_spin/lattice/pulse\_chunking.h}
as the \texttt{classical\_spin\_pulse\_chunking::build\_pulse\_segments}
free function so that all three lattice families can share it.
The window factor ($6\sigma$) and the free-segment $dt$ scale ($20
\times T_{\rm step}$) are exposed as
\texttt{kPulseWindowSigmas} / \texttt{kFreeDtFactor} constants and
have been chosen conservatively; a finer setting could squeeze out
a further $\sim 1.5 \times$ in the very-long $T_{\rm end}$ regime
but at the cost of robustness to atypical pulse shapes.

W3 is opt-in via \texttt{pulse\_window\_chunking} (default
\texttt{true}) in \texttt{SpinConfig}.  Setting it to \texttt{false}
falls back to the original single \texttt{integrate\_ode\_system}
call.

\subsection{Estimated speedup}

W1, W2 and W3 compose multiplicatively to leading order.  For a
typical $N_\tau = 100$ scan on a 16-core node with the spin
sector stationary, the expected speedup over the pre-W1/W2/W3
driver is
\[
\frac{2 N_\tau + 1}{N_\tau + 1} \cdot \min(N_\tau, T_{\rm threads})
\cdot s_{\rm W3}
\;\approx\;
2 \cdot \min(100, 16) \cdot 1.5
\;\sim\;
\mathbf{45 \times}
\]
in the regime where the inner RHS does not dominate (cheap
strain-bilin / coupled-spin RHS).  W1 contributes a factor of
$\sim 2$ (eliminating the $M_1$ branch), W2 fills all cores, and
W3 contributes the $\sim 1.5\times$ from the larger free-segment
$dt$.  For the heavy lattices where the inner RHS scales well to
all threads anyway, the W2 contribution shrinks toward $1$ and the
overall win is closer to $\sim 3 \times$, dominated by W1$+$W3.

The same three speedups apply per-rank in the MPI driver
(\texttt{pump\_probe\_spectroscopy\_mpi}); W1 broadcasts the M0
trajectory from rank 0 once and all other ranks synthesise their
local $M_1(\tau)$ values from it.  W2 is a no-op there because
$N_\tau$ is already distributed across MPI ranks.

\subsection{Why these are not done elsewhere}

Vampire and UppASD's pump--probe drivers (where they exist) do not
implement W1 because they are routinely run in the finite-$T$
Langevin regime where the time-shift identity does not hold.  Our
guard makes W1 safe to enable by default precisely because we
\emph{check} stationarity at runtime instead of assuming it.

W2 is straightforward in any code that treats lattice instances
as cheap to clone, but \texttt{Lattice}'s prior copy constructor
silently dropped the time-dependent drive parameters; this is
exactly the kind of bug that makes ``parallelise the outer loop''
look harder than it is.  Auditing and fixing the copy constructor
is the actual work of W2; the OpenMP region itself is six lines.

W3 is generic: any adaptive controlled stepper (Boost.Odeint,
\texttt{scipy.solve\_ivp} with \texttt{method='RK45'},
\textsc{SUNDIALS} \texttt{CVODE}, etc.) benefits from a larger
initial $dt$ when the dynamics is smooth.  We have not seen it
applied to 2DCS in the published spin codes; it is the cheapest
of the three to add and the most portable.

% =================================================================
\section{Ingredient XVI: packed row-major interaction tables and
         specialized inner kernels for the mixed-lattice MD RHS
         (M1--M4)}
\label{sec:packed-mixed-rhs}
% =================================================================

\texttt{MixedLattice}'s molecular-dynamics RHS
(\texttt{landau\_lifshitz}, called by Boost.Odeint inside RK4 / RK45
steps and ultimately by every $M_0$ and $M_1$ trajectory of the 2DCS
driver of Section~\ref{sec:pump-probe-w123}) is the throughput
bottleneck for any TmFeO$_3$-class run with the trilinear coupling
turned on.  Four small refactors (M1--M4) take it from $\sim 2.2 \times
10^3$ to $\sim 6.1 \times 10^3$ RHS/s at one thread on the same 16-core
benchmark socket --- a $2.8\times$ single-thread speedup on the
\texttt{tmfeo3-trilinear} model --- without changing any physics and
without touching the Monte-Carlo code path.

\subsection{The starting picture}

Before this pass, \texttt{get\_local\_field\_SU2\_flat\_into} already
wrote into a stack \texttt{double[3]} buffer (Ingredient~XIV);
likewise SU(3) wrote into \texttt{double[8]}.  The drive envelope was
already hoisted (Ingredient~VII).  But the inner trilinear loop was
literally
\begin{verbatim}
const auto& T = ti_SU2[n];                        // vector<MatrixXd>
for (size_t a = 0; a < spin_dim_SU2; ++a) {
    double temp = 0.0;
    for (size_t b = 0; b < spin_dim_SU2; ++b)
        for (size_t c = 0; c < spin_dim_SU2; ++c)
            temp += T[a](b, c) * state[p1+b] * state[p2+c];
    H[a] += temp;
}
\end{verbatim}
which has \emph{three} compounding defects:
\begin{enumerate}[noitemsep]
  \item \textbf{Pointer-chase per $a$.}  \texttt{SpinTensor3} is
        \texttt{std::vector<Eigen::MatrixXd>}, so \texttt{T[a]} is a
        \texttt{vector::operator[]} followed by an indirect load of
        \texttt{MatrixXd::m\_data}.
  \item \textbf{Wrong-stride access.}  \texttt{Eigen::MatrixXd} is
        column-major by default, but the innermost loop strides $c$
        fastest.  Each \texttt{T[a](b,c)} access therefore strides by
        \texttt{rows} doubles ($3$ or $8$) instead of unit stride,
        defeating prefetch and SIMD load coalescing.
  \item \textbf{Dimension is dynamic.}  The inner triple loop has
        runtime trip counts and cannot be unrolled or autovectorized
        by the compiler.
\end{enumerate}
The exact same three defects affect the bilinear branch.  And the
\texttt{landau\_lifshitz} dispatch opens \emph{two} OpenMP parallel
regions per call --- one over SU(2) sites, one over SU(3) sites.

\subsection{M1 + M4: pack everything into row-major \texttt{double[]}}

We add eight new per-site flat buffers to \texttt{MixedLattice}:
\begin{verbatim}
vector<vector<double>> bilinear_packed_SU2;       // (a,b) row-major
vector<vector<double>> bilinear_packed_SU3;
vector<vector<double>> mixed_bilinear_packed_SU2;
vector<vector<double>> mixed_bilinear_packed_SU3;
vector<vector<double>> trilinear_packed_SU2;      // (a,b,c) row-major
vector<vector<double>> trilinear_packed_SU3;
vector<vector<double>> mixed_trilinear_packed_SU2;
vector<vector<double>> mixed_trilinear_packed_SU3;
\end{verbatim}
The layout per site is
\[
  T^{n}_{abc}
  \;=\;
  \texttt{trilinear\_packed[site]}
  \bigl[
    n \cdot d_a d_b d_c
    + (a \cdot d_b + b) \cdot d_c
    + c
  \bigr]
\]
i.e.~contiguous and row-major in $(a,b,c)$.  These buffers are built
once at the end of the constructor, after the original
\texttt{vector<vector<SpinTensor3>>} graph has been assembled, by a
new \texttt{build\_packed\_interaction\_buffers()} method.  The
original storage is kept untouched for the Monte-Carlo path
(\texttt{site\_energy\_diff}, \texttt{total\_energy}, \dots) so M1 +
M4 are a pure addition.

\subsection{M2: compile-time-sized inner kernels}

With the data packed, the inner kernels become
\begin{verbatim}
template<size_t Da, size_t Db, size_t Dc>
inline void trilinear_kernel(const double* __restrict T,
                             const double* __restrict s1,
                             const double* __restrict s2,
                             double* __restrict H) {
    for (size_t a = 0; a < Da; ++a) {
        double acc = 0.0;
        for (size_t b = 0; b < Db; ++b) {
            const double s1b = s1[b];
            const double* Tab = T + (a*Db + b)*Dc;
            for (size_t c = 0; c < Dc; ++c) acc += Tab[c]*s1b*s2[c];
        }
        H[a] += acc;
    }
}
\end{verbatim}
and the bilinear analog \texttt{bilinear\_kernel<Da, Db>} takes a
single neighbour spin.  The hot path
($\texttt{spin\_dim\_SU2} = 3$, $\texttt{spin\_dim\_SU3} = 8$)
instantiates the compiler-known cases
\texttt{<3,3,3>} (27 FMAs), \texttt{<3,3,8>} (72 FMAs),
\texttt{<8,3,3>} (72 FMAs), \texttt{<8,8,8>} (512 FMAs);
GCC at \texttt{-O3 -march=native} fully unrolls these into AVX2 FMA
chains.  A runtime-dim fallback (\texttt{trilinear\_kernel\_dyn},
\texttt{bilinear\_kernel\_dyn}) preserves correctness for
non-standard SU($N$) dimensions.

\subsection{M3: single OpenMP team per RHS}

\texttt{landau\_lifshitz} now wraps both site loops in a single
\texttt{\#pragma omp parallel} region:
\begin{verbatim}
const size_t total_sites = lattice_size_SU2 + lattice_size_SU3;
#pragma omp parallel if(total_sites >= 64)
{
    #pragma omp for schedule(static) nowait
    for (size_t site = 0; site < lattice_size_SU2; ++site) { ... }
    #pragma omp for schedule(static)
    for (size_t site = 0; site < lattice_size_SU3; ++site) { ... }
}
\end{verbatim}
The \texttt{nowait} on the SU(2) loop lets fast threads start SU(3)
work while stragglers finish SU(2).  Each RHS now pays for one team
start-up + one barrier instead of two of each.  The fallback for
non-standard SU($N$) shapes still uses two separate parallel regions.

\subsection{Measured impact}

Re-running \texttt{bench\_md} on the same 16-core x86 socket as
Section~\ref{sec:bench-md-rhs-heap-free} (GCC 11
\texttt{-O3 -march=native -flto}, three repeats per cell):

\begin{center}
\small
\begin{tabular}{llrrrrr}
\hline
Family & Model ($L$) & $T_1$ (RHS/s) & $T_1$ ratio & $T_{16}$ (RHS/s) & $T_{16}$ ratio & $S(16)$ \\
\hline
\texttt{mixed} & TmFeO$_3$ trilinear ($L=4$) & $2.19 \times 10^{3} \to 6.10 \times 10^{3}$ & $\mathbf{2.79\times}$ & $1.96 \times 10^{4} \to 4.27 \times 10^{4}$ & $\mathbf{2.18\times}$ & $7.0\times$ \\
\texttt{mixed} & TmFeO$_3$ bilinear  ($L=4$) & --- $\to 7.07 \times 10^{3}$ & ---                & --- $\to 4.83 \times 10^{4}$ & ---                & $6.8\times$ \\
\texttt{mixed} & TmFeO$_3$ driven    ($L=4$) & --- $\to 7.03 \times 10^{3}$ & ---                & --- $\to 4.67 \times 10^{4}$ & ---                & $6.6\times$ \\
\texttt{mixed} & TmFeO$_3$ trilinear ($L=8$) & --- $\to 3.49 \times 10^{2}$ & ---                & --- $\to 3.50 \times 10^{3}$ & ---                & $10.0\times$ \\
\hline
\end{tabular}
\end{center}

The trilinear single-thread number went from $2.19 \times 10^{3}$ to
$6.10 \times 10^{3}$ RHS/s --- a $\mathbf{2.79\times}$ speedup.  The
strong-scaling factor at 16 threads dropped from $8.9\times$ to
$7.0\times$, which is the expected consequence of making the per-RHS
work cheaper (Amdahl): the OpenMP overhead is now a larger fraction
of the per-call budget on the small $L=4$ benchmark.  At $L=8$ the
team-management cost is amortized over more sites and the scaling
recovers to $10\times$ on 16 cores.  The bilinear and driven models
both gain similarly (the M4 packing flips the same column-major bug
in their inner loop) and now run at $\sim 7 \times 10^{3}$ RHS/s
single-threaded --- about $1.5\times$ over the old common case.

\subsection{Why this win was sitting on the table}

The Monte-Carlo $\Delta E$ path (\texttt{site\_energy\_*\_diff})
already used the natural ``pre-contract partners then dot with
\texttt{spin\_diff}'' factorisation (item~14 of the checklist), but
that pre-contraction lives in MC code and is a fundamentally
different numerical trick: it factors the $\Delta E$ formula
algebraically.  The MD RHS does the \emph{same arithmetic} as before
M1--M4, just on a memory layout that the compiler can vectorize.
The MC and MD paths share the original
\texttt{vector<vector<SpinTensor3>>} as the source of truth for the
coupling tensors; M1 + M4 add a parallel SoA copy used only by the
MD hot path.

There is no analogous packed-trilinear table in
\texttt{Lattice::get\_local\_field\_flat} either.  That code already
ships the packed-bilinear (\texttt{bi\_flat\_J} + offsets, item~3 of
the checklist) but its trilinear branch still runs the
\texttt{T[a](b,c)} pattern with the same column-major access bug; a
follow-up pass can apply the same M1+M2 treatment there.  This is
deferred to keep the present change scoped to the
\texttt{MixedLattice} ODE that the user asked about.

% =================================================================
\section{Putting it together: a checklist}
% =================================================================

When auditing a classical-spin MC code for ``is this SOTA?'', go down
this list.  Each item is independent and the table-stakes ingredient
that any serious atomistic spin code (Vampire, UppASD, Spirit) ships
with.

\begin{enumerate}[itemsep=2pt]
  \item \textbf{Inner loop is allocation-free.}  No \texttt{vector<>},
        no \texttt{Eigen::VectorXd}, no \texttt{new}, no \texttt{shared\_ptr}
        inside the per-site work.
  \item \textbf{Spins are stored contiguously.}  Iterating over sites is
        a linear memory scan; neighbours are accessed via a precomputed
        index table.
  \item \textbf{Interaction tables are flat.}  Bilinear $\mathsf{J}$
        matrices live in one \texttt{double[]} of size $N \times z \times 9$.
  \item \textbf{Sweep order is sequential} (or, where required, a static
        checkerboard---never per-call random with replacement).
  \item \textbf{Compiler flags are aggressive.}  \texttt{-O3
        -march=native -flto -ffp-contract=fast}, at minimum.
  \item \textbf{Acceptance test short-circuits.}  Use \texttt{||}, not
        \texttt{|}, and put the cheap test first.
  \item \textbf{Cluster algorithms are available.}  Wolff and Swendsen--Wang
        for $\mathbb{Z}_2$-symmetric problems; embedded-cluster
        variants for continuous spins.
  \item \textbf{Overrelaxation is interleaved with Metropolis} for
        continuous spins.
  \item \textbf{Parallel tempering uses persistent buffers and
        \texttt{Sendrecv\_replace}.}
  \item \textbf{The temperature ladder is feedback-optimized.}
  \item \textbf{Per-rank, per-thread RNGs are deterministically seeded.}
  \item \textbf{Observables accumulators precompute lookup tables} once
        in \texttt{initialize()}, never in the per-sample call.
  \item \textbf{The molecular-dynamics integrator can be re-projected
        onto the constraint surface} (here, $|\mathbf{S}| = 1$) at user
        request.
  \item \textbf{Trilinear / multi-spin terms pre-contract their partner
        indices} (Section~\ref{sec:trilinear}) so that one Metropolis
        proposal costs $O(d^{3}) + O(d)$ instead of $2\,O(d^{3})$.
  \item \textbf{Linear combinations of derivative operators are fused
        into a single neighbour pass} (Section~\ref{sec:me-fuse}),
        which is the dominant cost in spin--lattice / magnetoelastic
        models such as the strain--phonon honeycomb here.
  \item \textbf{Sparse Lie-algebra structure constants are written
        as a hand-rolled non-zero list, not a dense triple loop}
        (Section~\ref{sec:sparse-su3}).  For SU(3) this turns the
        $8 \times 8 \times 8 = 512$ multiply--add tensor contraction
        into nine fused-multiply--adds.  The same trick (with the
        Levi--Civita symbol) is implicit in the SU(2) cross product;
        the SU(N) generalisation is the same statement.
  \item \textbf{Time-dependent envelopes are hoisted out of the
        per-site RHS loop, with an early exit when the drive is
        off} (Section~\ref{sec:hoist-time}).  This includes the
        elimination of \texttt{Eigen::VectorXd} returns from
        \texttt{get\_local\_field\_*} in favour of caller-supplied
        \texttt{double*} buffers, and the elimination of the
        save-then-restore copy of \texttt{cell\_strains\_} in the
        strain RHS (Section~\ref{sec:strain-rhs}).
  \item \textbf{Intra-sweep parallelism uses a static sublattice
        colouring} of the bond graph, not a hash-based or
        ``parallel-with-locks'' scheme
        (Section~\ref{sec:sublattice-colour}).  Sites of the same
        colour are provably independent, so the parallel kernel is
        an exact Metropolis sweep at any thread count, with the
        same per-thread \texttt{thread\_local} Lehmer RNG used by
        the serial path.
  \item \textbf{The molecular-dynamics RHS opens at most one
        OpenMP team per call} (Section~\ref{sec:rhs-heap-free}).
        \texttt{ode\_system} wraps the spin sync, the spin EOM,
        and the serial preamble/epilogue in a single
        \texttt{\#pragma omp parallel} with \texttt{nowait} +
        \texttt{single} blocks for the serial pieces, so a typical
        RHS pays for one team start-up and two implicit barriers
        instead of two complete team start-ups.  The per-site loop
        body itself is heap-free: a stack
        \texttt{Eigen::Vector3d} accumulator,
        \texttt{Map<const Matrix3d>}-based exchange products with
        \texttt{noalias()}, and \texttt{compute\_*\_v3} ME / ring-
        exchange helpers that fuse the eight Eg sub-derivatives
        into one neighbour pass.  This is what removed the
        $\sim 11$ \texttt{malloc/free} pairs per site per RHS that
        had been capping the strain--phonon Langevin integrator.
  \item \textbf{The 2DCS / pump--probe driver collapses redundant
        integrations and parallelises the $\tau$ loop}
        (Section~\ref{sec:pump-probe-w123}).  Three independent
        speedups apply to every \texttt{pump\_probe\_spectroscopy}
        entry point:  W1 synthesises $M_1(\tau)$ from the M0
        trajectory by time-shift (guarded at runtime by a
        $\max_i |\dot S_i|_\infty$ stationarity check, so it is
        silently disabled if the user has not actually equilibrated
        the spin sector or is running finite-$T$ Langevin
        dynamics);  W2 distributes the $\tau$ loop across OpenMP
        threads using deep-cloned lattice instances (this is the
        reason the \texttt{Lattice} copy constructor must include
        the time-dependent drive parameters and the AFM sublattice
        signs);  W3 splits the integration into pulse-active and
        free segments and feeds the controlled stepper a
        $\sim 20\times$ larger initial $dt$ in the free segments,
        letting the adaptive RK grow its step rapidly while the
        observer cadence stays at \texttt{T\_step}.  The shared
        helper lives in
        \texttt{include/classical\_spin/lattice/pulse\_chunking.h}
        so all three lattice families
        (\texttt{Lattice}, \texttt{MixedLattice},
        \texttt{PhononLattice}) use the same pulse-window logic.
  \item \textbf{Bilinear and trilinear interaction tensors are
        packed row-major in the MD RHS and consumed by
        compile-time-sized inner kernels}
        (Section~\ref{sec:packed-mixed-rhs}).  The Eigen
        \texttt{MatrixXd} default storage is column-major and the
        natural inner loop strides $c$ fastest, so
        \texttt{T[a](b,c) * S[b] * S[c]} pays both a
        \texttt{vector<MatrixXd>} pointer chase per $a$ and a
        non-unit stride per $c$.  We pack the trilinear and
        bilinear couplings (and their mixed SU(2)--SU(3) partners)
        into a single \texttt{double[]} per site at the end of
        the constructor and dispatch through templated
        \texttt{trilinear\_kernel<Da,Db,Dc>} /
        \texttt{bilinear\_kernel<Da,Db>} kernels for the canonical
        $3\times3\times3$, $3\times3\times8$, $8\times3\times3$,
        $8\times8\times8$, $3\times3$, $3\times8$, $8\times3$,
        $8\times8$ shapes; the compiler unrolls these into AVX2
        FMA chains.  Combined with merging the SU(2) and SU(3)
        parallel regions of \texttt{landau\_lifshitz} into a
        single \texttt{\#pragma omp parallel} (item~18 already
        applies to the strain RHS; this extends it to mixed),
        TmFeO$_3$ trilinear MD throughput goes from
        $2.2 \times 10^{3}$ to $6.1 \times 10^{3}$ RHS/s on one
        thread ($\mathbf{2.8\times}$).
\end{enumerate}

\texttt{ClassicalSpin\_Cpp} now ticks all twenty, including
item~3 (the Structure-of-Arrays bilinear table, see
Section~\ref{sec:soa}), the Marsaglia sphere sampler
(Section~\ref{sec:marsaglia}, including its mixed/strain
extensions), the trilinear pre-contraction of
Section~\ref{sec:trilinear}, the magnetoelastic-derivative
fusion of Section~\ref{sec:me-fuse}, the sparse SU(3)
structure-constant kernel of Section~\ref{sec:sparse-su3}, the
drive-envelope hoist + heap-free local field of
Section~\ref{sec:hoist-time}, the sublattice-coloured intra-sweep
OpenMP parallelism of Section~\ref{sec:sublattice-colour},
and---new in this pass---the persistent OpenMP region with the
heap-free \texttt{compute\_*\_v3} family of strain-phonon RHS
helpers (Section~\ref{sec:rhs-heap-free}), which closed the
$\sim 21\times$ single-thread gap to the lattice/mixed RHS
throughput on the local-strain code path.  The remaining gap before
head-to-head benchmarking against Vampire/UppASD on large
Heisenberg lattices ($L \gtrsim 128$) is hand-rolled SIMD
intrinsics on the SoA bilinear table---the layout is now
AVX2-compatible (9-double blocks for the $J$ matrix), but we still
rely on the compiler's auto-vectorizer to emit
\texttt{vfmadd231pd}.  An explicit intrinsics rewrite would be
expected to give an additional $1.5$--$2 \times$ on the per-site
energy difference, but the cost in maintainability and portability
is real.

% =================================================================
\section{Ingredient XVII: final audit of the MixedLattice 2DCS
         pipeline (W1 / W2 / W3 / M1--M4) and a corrected
         W1$\,$+$\,$\texttt{save\_spin\_trajectories} interaction
         in the MPI driver}
% =================================================================
\label{sec:audit-mixed-2dcs}

After landing the inter-$\tau$ parallelism (W2), the time-shift
synthesis of $M_1(\tau)$ from $M_0$ (W1), the pulse-window-aware
integration chunking (W3), the row-major packed bilinear / trilinear
tables (M1, M4), the templated inner kernels (M2), and the
single-team \texttt{landau\_lifshitz} (M3), we did a
file-by-file audit of the full \texttt{MixedLattice} 2DCS pipeline.
The audit covered the W1/W2/W3 wiring, the drives
(\texttt{single\_pulse\_drive} / \texttt{double\_pulse\_drive}),
observable accumulation, the HDF5 streaming writer, the per-rank
$\tau$ distribution, the M0 broadcast, all heap allocations on the
hot path, and the integrator settings (\texttt{dopri5} in
\texttt{integrate\_const} mode at $10^{-10}$).  Two findings were
material; the rest were either documented design choices or
sub-percent niceties that we leave for a future pass.

\paragraph{Finding 1 (\textbf{correctness fix}).}  The MPI driver
\texttt{pump\_probe\_spectroscopy\_mpi} had a silent data-corruption
bug when both \texttt{reuse\_m0\_for\_m1} and
\texttt{save\_spin\_trajectories} were on.  The W1 synthesis path
returns only the magnetisation triples
$(M_{\mathrm{stag}}, M_{\mathrm{loc}}, M_{\mathrm{glob}})$ at every
sample, not the per-site spin field, so the worker's
\texttt{local\_spin\_flat\_M1} buffer was empty for every $\tau$.
The MPI send used \texttt{sf\_m1.size() = 0} doubles, but
rank~0's \texttt{MPI\_Recv} requested
\texttt{state\_traj\_size\_spins = time\_points * state\_dim}
doubles---a tag-matched, count-mismatched message.  Most MPI
implementations satisfy a count-mismatched recv as soon as the
matching tag arrives and report the actual byte count via
\texttt{MPI\_Status}; the user-supplied buffer past the actual
sent length keeps its \emph{previous} contents.  Those contents
were then written verbatim into the HDF5 file as the M1 spin
trajectory.  The bug was latent because the default 2DCS run does
\emph{not} request \texttt{save\_spin\_trajectories}; only users
who turned that flag on plus relied on W1 (the default) saw it.

The fix lives in \texttt{src/core/mixed\_lattice\_md.cpp} and is
deliberately minimal: when the user requests
\texttt{save\_spin\_trajectories}, rank~0 forces
\texttt{can\_reuse\_m0\_flag = 0} \emph{before} the broadcast, and
emits an explicit log line explaining that W1 is being skipped
because the synthesis cannot reproduce per-site states.  All ranks
then take the integrate-from-scratch path, the per-$\tau$
\texttt{M1\_spin\_states} is populated by
\texttt{single\_pulse\_drive}, the worker MPI send carries the full
\texttt{time\_points * state\_dim} doubles, and the HDF5 file is
correct.  As a defensive belt-and-braces measure, the empty-buffer
fall-through (which can still trigger only if the integrator
returns zero samples for some pathological window) now pads with
\texttt{NaN} instead of \texttt{vector::emplace\_back()} so any
downstream consumer immediately sees the gap.  The serial driver
\texttt{pump\_probe\_spectroscopy} does not accept
\texttt{save\_spin\_trajectories} and is unaffected.

\paragraph{Finding 2 (documented limitation).}  W3 is currently
CPU-only.  The GPU drives (\texttt{single\_pulse\_drive\_gpu},
\texttt{double\_pulse\_drive\_gpu}) ignore the
\texttt{pulse\_window\_chunking} flag and integrate the full
\texttt{[T\_start, T\_end]} window in one segment.  This is
intentional---chunking the GPU integrator currently requires
re-issuing kernels per segment, which hurts more than the larger
free-segment $dt$ helps---but should be revisited when GPU spin
counts grow into the regime where the controlled stepper would
naturally take much larger steps in the free window.  The CPU
fallback (which is what 2DCS production runs use today) retains
the full W1$\,$+$\,$W2$\,$+$\,$W3 stack.

\paragraph{Combined effect on a 2DCS \texttt{MixedLattice} run.}
For a TmFeO$_3$ trilinear setup with $L = 8$, $\tau$-grid of
$N_\tau = 64$ points, $T_{\mathrm{end}} - T_{\mathrm{start}} =
20$\,ps and \texttt{dopri5} adaptive stepping, the cumulative
speedup over a vanilla single-threaded driver-without-W1/W2/W3 is
$\approx 32 \times$ on a 16-core node (W2 contributes
$\approx 11\times$, W3 $\approx 1.6\times$, W1 $\approx 1.7\times$
when the spin sector is genuinely stationary, M1--M4
$\approx 2.8\times$ at the RHS level which propagates to the same
factor at the spectroscopy level since the integrator is RHS-bound).
Users who want a sanity check that the audit fixes are in place
should grep for ``\texttt{[W1] Disabled because
save\_spin\_trajectories=true}'' in the MPI driver's stdout when
running with both flags on.

\paragraph{Items deliberately left for a future pass.}
\begin{itemize}[noitemsep]
  \item Per-$\tau$ HDF5 writes are serialised under the W2
        \texttt{\#pragma omp critical(mixed\_pump\_probe\_hdf5\_write)}.
        For a $\tau$-grid where each integration takes seconds
        and each HDF5 chunk write is sub-millisecond this is
        invisible; on extremely lightweight problems where the
        integration itself takes only milliseconds the critical
        section will throttle W2.  The fix would be a per-thread
        in-memory buffer + a single-threaded flush phase, or
        delegating I/O to a dedicated rank.  Not done now because
        the regime where this matters is already faster than the
        physics requires.
  \item The drive observer's per-sample
        \texttt{Eigen::Map}\,/\,\texttt{SpinVector}\,/\,
        \texttt{push\_back} cycle creates short-lived
        \texttt{Eigen::VectorXd} temporaries.  These do not
        currently allocate on the heap (they are
        $\le 8$-element vectors and Eigen falls back to stack
        storage at small fixed sizes), but a future
        \texttt{std::array}-based trajectory representation would
        eliminate them entirely.
  \item \texttt{MixedLattice} relies on the implicit copy
        constructor for W2 cloning.  Today every member is
        value-typed (\texttt{std::vector}, \texttt{Eigen} dense
        matrices, plain doubles) so the implicit copy is correct
        and complete.  An explicit copy constructor would
        document the contract and prevent a future pointer-typed
        member from silently breaking W2.  The runtime risk is
        zero today; the maintenance risk is real and noted here.
\end{itemize}

% =================================================================
\section{What the numbers mean}
% =================================================================

The headline number from Table~\ref{tab:bench} is
$\sim 1.8 \times 10^{7}$ Metropolis spin-updates/sec/core for a
short-range Heisenberg model on a single x86 core (up from
$\sim 1.4 \times 10^{7}$ before this pass).  For comparison,
\textsc{Vampire}'s reported figures on Skylake-class CPUs are in the
range $10^{7}$--$10^{8}$ spin-updates/sec/core depending on the model;
\textsc{UppASD} reports $\sim 5 \times 10^{7}$ for similar problem
sizes.  We are within a factor of $\sim 2$--$3$ of those codes
\emph{without} hand-rolled SIMD intrinsics.  Closing the rest of the
gap requires explicit AVX2/AVX-512 work on the per-site energy
difference; the SoA layout (Section~\ref{sec:soa}) is the necessary
prerequisite for that, and it is now in place.

With the sublattice-coloured intra-sweep parallelism of
Section~\ref{sec:sublattice-colour}, the same Heisenberg-honeycomb
$L = 32$ kernel reaches $\sim 1.1 \times 10^{8}$ Metropolis
updates/sec on a single 16-core socket
($T_1 \approx 0.25$~s, $S(16) \approx 6.6$, throughput $\approx
6.6 \times 1.6 \times 10^{7}\,\text{u/s}$); the overrelaxation kernel
reaches $\sim 2.0 \times 10^{8}$ updates/sec.  This puts a
single-node \texttt{ClassicalSpin\_Cpp} run squarely in the
upper end of the Vampire/UppASD per-node throughput range, before
any explicit SIMD intrinsics work.

The cache-hierarchy effect is worth noting on its own.  At $L = 32$ the
spin data ($N \times d \times 8 = 49$~kB) fits in L1; at $L = 64$
(196~kB) it spills to L2; at $L = 128$ (786~kB) it spills to L3.  The
measured throughput drops from $1.8 \times 10^{7}$ to $6.5 \times 10^{6}$
across this range---a $\sim 3 \times$ slowdown that is essentially
\emph{all memory bandwidth}.  This is the regime where SIMD and the
already-flat SoA layout would pay the most, and where the
``flat-spin scratch'' refactor we deliberately did \emph{not}
land at $L = 32$ (Section~\ref{sec:marsaglia}, last paragraph) is
likely to start winning.

% =================================================================
\section{Ingredient XVIII: integrator-layer audit --- dispatch
correctness and tolerance defaults}
% =================================================================
\label{sec:ing-xviii}

The XV/XVI/XVII series targeted everything \emph{around} the ODE
integrator: the magnetisation symmetry that lets us synthesise
$M_1(\tau)$ from a single $M_0$ run (W1), the inter-$\tau$ OpenMP
parallelism (W2), the pulse-window-aware time stepping (W3), the
packed interaction tables (M1/M4), and the persistent OpenMP region
in the RHS (M3).  This section turns the spotlight on the integrator
\emph{itself}.  We found two issues, both with the same character:
the library was silently doing more work per integration step than
the physics required.

\subsection*{XVIII-A.  Pump--probe default tolerance was 100$\times$
too tight}

Every \texttt{*\_pulse\_drive()} entry point across
\texttt{Lattice}, \texttt{MixedLattice}, and \texttt{PhononLattice}
called \texttt{integrate\_ode\_system(\dots, abs\_tol = 1e-10,
rel\_tol = 1e-10)}.  That tolerance was hard-coded as a literal in
twelve different places and was inherited from an early
exploratory commit, not a calibration against the 2DCS observable.

The physics is forgiving.  A typical $\chi^{(3)}$ feature is
$\sim 10^{-3}$ of $|\mathbf{M}|$, so $\sim 5$ significant figures
in the trajectory ($\approx 10^{-8}$ absolute on a normalised
spin) is more than enough.  For Dormand--Prince 5(4) the accepted
step size scales as $\Delta t \propto \text{tol}^{1/5}$, so
relaxing tol from $10^{-10}$ to $10^{-8}$ should give
$(10^{2})^{1/5} \approx 2.5\times$ larger steps and a matching
$\sim 2.5\times$ fewer RHS calls per integration when the
controller is tolerance-limited.

Concretely we:
\begin{itemize}
  \item Added two named constants
        \texttt{kDefaultPumpProbeAbsTol = kDefaultPumpProbeRelTol
        = 1e-8} to \texttt{include/classical\_spin/lattice/pulse\_chunking.h}.
  \item Threaded \texttt{abs\_tol} / \texttt{rel\_tol} parameters
        (defaulting to the new constants) through every
        \texttt{single\_pulse\_drive}, \texttt{double\_pulse\_drive},
        \texttt{pump\_probe\_spectroscopy}, and
        \texttt{pump\_probe\_spectroscopy\_mpi} signature in all
        three lattice families, and replaced the twelve hard-coded
        \texttt{1e-10} call-site literals with the new parameters.
  \item Added a parallel \texttt{md\_abs\_tol} / \texttt{md\_rel\_tol}
        pair (default \texttt{1e-6}) and
        \texttt{pump\_probe\_abs\_tol} / \texttt{pump\_probe\_rel\_tol}
        pair (default \texttt{1e-8}) to \texttt{SpinConfig}, plumbed
        them through every runner, and made the \texttt{molecular\_dynamics}
        path use a $-1.0$ sentinel that falls back to the
        method-aware defaults when the user does not override.
  \item Pinned the change with a regression test
        (\texttt{test\_pulse\_drive\_tolerance}) that runs
        \texttt{single\_pulse\_drive} on a small TmFeO$_3$ mixed
        lattice at \texttt{1e-10} and \texttt{1e-8} and asserts
        the worst $|\Delta M|$ across all six magnetisation triples
        and all sample times stays under $5 \times 10^{-6}$.
\end{itemize}

\paragraph{Measured speedup (from \texttt{bench\_integrators}).}
On a $1\times1\times1$ TmFeO$_3$ lattice with a Gaussian pulse
($\sigma = 0.2$, $T \in [-1, 8]$) the wall-clock time per
\texttt{single\_pulse\_drive} call (averaged over five runs):
\begin{center}
\begin{tabular}{lcc}
\hline
\textbf{method} & \textbf{tol} & \textbf{wall (ms)} \\
\hline
dopri5 & $10^{-8}$  & $\sim 30$ \\
dopri5 & $10^{-10}$ & $\sim 71$ \\
\hline
\end{tabular}
\end{center}
That is $\mathbf{2.4\times}$ speedup, almost exactly the
theoretical $2.5\times$ bound.  At $L = 2$ ($64$ mixed sites) the
controller is no longer tolerance-limited (the pulse-window
chunking and other constraints dominate $\Delta t$ selection) and
the same comparison gives $\sim 1.0\times$ --- i.e.\ the relaxation
is \emph{free} on small systems and \emph{neutral} on larger ones.
There is no regime in which it is a regression.

\subsection*{XVIII-B.  \texttt{rk54}/\texttt{rkf54} was misrouted
to a 7(8) integrator}

The user-facing integrator name \texttt{rk54} (and its alias
\texttt{rkf54}) is documented as ``Runge--Kutta 5(4) with embedded
4th-order error estimate''.  In both \texttt{Lattice} and
\texttt{MixedLattice} the dispatch table actually instantiated
\texttt{boost::numeric::odeint::runge\_kutta\_fehlberg78}, a
\emph{7(8)} method that uses 13 RHS evaluations per accepted step
versus the 6 of a true 5(4).  This is a pure correctness bug at
the API level: a user asking for a 5(4) method got a 7(8) method
silently.

The fix is a one-line change in each header: route
\texttt{rk54}/\texttt{rkf54} to
\texttt{runge\_kutta\_cash\_karp54}, which is the closest 5(4)
method available in \texttt{boost::numeric::odeint} (Dormand--Prince
is also 5(4) but is exposed under the separate name
\texttt{dopri5}).  \texttt{PhononLattice} was already correct, so
this only touched the two SU(2)/SU(2)+SU(3) headers.

\paragraph{Performance side-effect.}  On the same
\texttt{bench\_integrators} micro-benchmark at $L = 1$ the
\emph{previous} (buggy) \texttt{rk54} dispatch was actually
$\sim 3\times$ \emph{faster} than the (now-correct) Cash--Karp
5(4), because Fehlberg 7(8) takes much larger steps on smooth
post-pulse dynamics.  This is not a reason to keep the bug:
\begin{itemize}
  \item Users who explicitly want the 7(8) method can now write
        \texttt{md\_integrator = "rkf78"}, which always meant 7(8).
  \item At $L = 2$ the (now-correct) Cash--Karp 5(4)
        \texttt{rk54} is \textbf{$\mathbf{1.7\times}$ faster than
        \texttt{dopri5}} at the same tolerance, so the
        correctness fix actually \emph{improves} the realistic
        large-system case for users who had explicitly chosen
        \texttt{rk54}.
\end{itemize}

\subsection*{XVIII-C.  Fixed-step \texttt{rk4} is not a free lunch}

When asked ``what about \texttt{rk4}?'' the natural hope is that
the cheapest classical integrator (4 RHS calls per step, no error
control) might out-perform the adaptive 5(4) family on a smooth
post-pulse evolution.  We measured this carefully on the same
TmFeO$_3$ benchmark (with the bench reporting trajectory error
against a \texttt{dopri5}@$10^{-12}$ reference) and the answer is
\textbf{no}:

\begin{center}
\small
\begin{tabular}{lccp{6cm}}
\hline
\textbf{method} & \textbf{dt} & \textbf{wall (ms)} & \textbf{$\max |M-M_{\text{ref}}|$ \quad / \quad NaN\_at} \\
\hline
rk4 & $0.05$    & $0.7$  & $\infty$, blows up at $t=-0.1$ (pre-pulse) \\
rk4 & $0.02$    & $1.5$  & $\infty$, blows up at $t=-0.18$ \\
rk4 & $0.01$    & $3.0$  & $\infty$, blows up at $t=-0.20$ \\
rk4 & $0.005$   & $5.8$  & $\sim 1$ (stable but 100\% error) \\
rk4 & $0.002$   & $14$   & $0.55$ (50\% error) \\
rk4 & $0.001$   & $29$   & $0.039$ (4\% error, still 4 orders worse than dopri5) \\
\hline
dopri5 & $10^{-8}$ & $29$ & $1.6 \times 10^{-6}$ (the new default) \\
\hline
\end{tabular}
\end{center}

The story the bench tells:

\begin{itemize}
  \item \emph{Stability.}  The TmFeO$_3$ Hamiltonian
        ($J_{1ab} = 4.74$, $J_{1c} = 5.15$ plus trilinear
        couplings $u, v, w \sim 0.1$--$0.3$, all in our energy units
        where time is dimensionless) has an effective spectral
        radius for the LLG Jacobian large enough that
        \texttt{rk4} requires $\Delta t \lesssim 0.005$ just to be
        \emph{numerically stable}.  Anything larger gives a
        silent NaN trajectory.
  \item \emph{Accuracy.}  To match the trajectory accuracy of
        \texttt{dopri5}@$10^{-8}$ (the new pump-probe default),
        \texttt{rk4} would need $\Delta t \lesssim 3 \times 10^{-4}$,
        i.e.\ $\sim 30\,000$ steps over $T = [-1, 8]$, which is
        $\sim 3 \times$ \emph{slower} than the adaptive default it is
        meant to replace.
  \item \emph{Safety.}  The fixed-step methods have no error
        control: if the user's \texttt{md\_timestep} is too large
        for the Hamiltonian's spectral radius, the result is a
        NaN-laden trajectory, not an error message.  The adaptive
        5(4) family at least flags this by aborting with the
        controller's ``step size too small'' diagnostic.
\end{itemize}

\paragraph{Recommendation.}  Use \texttt{rk4} only for code paths
where $\Delta t$ is externally constrained (e.g.\ coupling to a
fixed-step phonon integrator), and \emph{always} cross-check the
\texttt{rk4} trajectory against the same problem run with
\texttt{dopri5}@$10^{-8}$.  For the standard pump-probe / 2DCS
workflow, the \texttt{dopri5} default already wins on every axis.

\paragraph{Caveat about \texttt{rkf78}.}  An earlier draft of this
section recommended \texttt{rkf78} as a $\sim 4\times$ speedup
based on wall-clock measurements of \texttt{single\_pulse\_drive}.
\textbf{That recommendation was wrong}: the bench was not checking
trajectory finiteness, so a NaN-laden trajectory looked ``fast''
because the 7(8) controller takes huge steps and finishes quickly.
Adding the \texttt{NaN\_at} column to the bench reveals that on
the small TmFeO$_3$ lattice \texttt{rkf78} blows up post-pulse
(at $t \approx 3.2$) at \emph{both} \texttt{1e-8} and
\texttt{1e-10} tolerances.  Until the \texttt{rkf78} controller
behaviour on the trilinear-coupling RHS is investigated separately,
do not recommend \texttt{rkf78} as a default speedup.  The honest
recommendation is:
\begin{verbatim}
md_integrator = dopri5      # the safe default
pump_probe_abs_tol = 1e-8   # already the new default
pump_probe_rel_tol = 1e-8
\end{verbatim}

\subsection*{Net effect}

For the canonical 2DCS workflow on a small mixed lattice the
changes together give a $\sim 2.4\times$ pump-probe speedup
(XVIII-A) with no observable change in the magnetisation trajectory
beyond $\sim 10^{-6}$ on $|\mathbf{M}|$, and silently fix the
\texttt{rk54}/\texttt{rkf54} misroute so users get the 5(4) method
they asked for (XVIII-B).  The 2DCS observable \emph{never depends}
on accuracy at this level: a $\chi^{(3)}$ feature is $\sim 10^{-3}$
of $|\mathbf{M}|$, so the $\sim 10^{-6}$ trajectory difference is
three orders of magnitude below the smallest physically meaningful
signal.

The audit also produced two negative results (XVIII-C and the
\texttt{rkf78} caveat) that we record explicitly because the
naive ``bench-by-wall-clock-only'' approach previously gave the
wrong answer for both.  The lesson generalises: any future
integrator-level audit \emph{must} measure trajectory error against
a high-accuracy reference, not just wall-clock time.

% =================================================================
\section{Ingredient XIX: the 2DCS reader pipeline ---
         FFT linearity, batched FFTs, gated debug PDFs}
% =================================================================

The C++ side of a 2DCS run produces an HDF5 file with a reference
trajectory and one perturbed-pair trajectory per $\tau$ delay
(see Ingredient XV).  The Python reader
\texttt{util/readers\_new/reader\_TmFeO3.py}, in particular
\texttt{read\_2DCS\_combined\_hdf5}, is responsible for
\begin{enumerate}[noitemsep,leftmargin=*]
  \item loading the per-component magnetisation streams
        $M_0$, $M_1$, $M_{01}$ from the HDF5 file,
  \item assembling the third-order observable
        $M_{NL}=M_{01}-M_0-M_1$,
  \item Fourier-transforming it on the $(\tau,t)$ grid,
  \item building several $\lambda$-mode mixtures (e.g.\
        $\lambda_5+\lambda_7$, $\lambda_2+\lambda_5$,
        $\lambda_2+\lambda_7+S_x$, \ldots) for the SU(3) sublattice,
  \item rendering up to thirty PDF figures.
\end{enumerate}
On the 34\,MB \texttt{nocouple\_ctrl} benchmark
($n_t=10\,000$, $n_\tau=6$, $\dim_{\mathrm{SU2}}=3$,
$\dim_{\mathrm{SU3}}=8$) this took \textbf{17\,s} per call
and was the dominant cost of any
parameter sweep that produces $O(10^2$--$10^3)$ HDF5 outputs.

\paragraph{What the profiler said.}
A surgical monkey-patch profile of \texttt{savefig},
\texttt{Axes.imshow}, and \texttt{h5py.Dataset.\_\_getitem\_\_}
revealed three problems, in order of severity:
\begin{enumerate}[noitemsep,leftmargin=*]
  \item All five $\lambda$-mode debug PDFs
        (\texttt{lambda57}, \texttt{lambda25}, \texttt{lambda257},
        \texttt{lambda27x}, \texttt{lambda27x\_v2}) were rendered
        \emph{regardless of} \texttt{skip\_plots=True} ---
        $\sim 75\%$ of total wall time.
  \item The six per-signal component PDFs
        (\texttt{M0/M1/M01\_SU2/SU3\_components\_debug.pdf},
         8 or 3 subplots each) were also unconditional.
  \item Each $\lambda$-mode magnitude spectrum was computed by
        building the time-domain combination
        $a\,X(\tau,t)+b\,Y(\tau,t)$ and then re-running
        \texttt{np.fft.fft2} --- thirty $4926\times6$ FFTs were
        executed per call, twenty of them redundant.
\end{enumerate}

\paragraph{Fixes.}
\begin{enumerate}[noitemsep,leftmargin=*]
  \item All debug-PDF blocks (lines around 1797, 2113, 2316, 2449,
        2578, 2708, 2837 in
        \texttt{reader\_TmFeO3.py}) were wrapped in
        \texttt{if not skip\_plots:}.  The underlying FFTs and
        on-disk \texttt{.npy} writes were lifted \emph{outside}
        the conditional, so the spectra are still produced when
        plotting is suppressed.
  \item The HDF5 load loop was vectorised: instead of an inner
        \texttt{for comp in range(spin\_dim)} that touches one
        component at a time, we transpose
        $(n_t,\dim)\to(\dim,n_t)$ once and slice
        \texttt{M\_NL[:, i, :n] = M01\_T - M0\_T - M1\_T}.
        The intermediate \texttt{M0\_comp/M1\_comp/M01\_comp}
        arrays (each $\dim\times n_\tau\times n_t$, up to
        $8\times 6\times 10^4$ doubles
        $= 3.7\,$MB) are now allocated only when
        \texttt{save\_intermediates=True}.
  \item A complex-FFT cache \texttt{\_cfft\_cache} was added
        alongside the existing magnitude cache.  All per-component
        FFTs are computed in a \emph{single} batched call
        \begin{verbatim}
        scipy.fft.fft2(data_3d, axes=(-2,-1), workers=-1)
        \end{verbatim}
        which amortises FFT-plan setup and uses every available
        core for the larger transforms.  The $\lambda$-mode
        mixtures are then synthesised in frequency space by
        linearity:
        \[
            \lvert \mathcal F\{aX+bY\}\rvert
            \;=\; \lvert a\,\hat X + b\,\hat Y \rvert ,
        \]
        i.e.\ one weighted sum and one $\lvert\cdot\rvert$ per
        derived spectrum, instead of one fft2.  This is
        implemented by \texttt{compute\_2d\_fft\_combo}, with a
        graceful fall-back to the time-domain path when
        a cache entry happens to be missing.
  \item The CLI gained \texttt{--skip-plots},
        \texttt{--no-save-intermediates}, and a
        \texttt{--fast} convenience flag (equivalent to both
        of the previous two).  These plumb through to the
        existing keyword arguments of
        \texttt{read\_2DCS\_combined\_hdf5}.
\end{enumerate}

\paragraph{Numbers.}
On the 34\,MB \texttt{nocouple\_ctrl} benchmark
($n_t=10\,000$, $n_\tau=6$, $\dim_{\mathrm{SU2}}=3$,
$\dim_{\mathrm{SU3}}=8$):
\begin{center}
\begin{tabular}{lrr}
\hline
\textbf{Mode} & \textbf{Before XIX} & \textbf{After XIX}\\
\hline
\texttt{--skip-plots --save-intermediates}    & 17.0\,s & 0.06\,s \\
\texttt{--skip-plots --no-save-intermediates} & 17.0\,s & 0.01\,s \\
\texttt{--fast} (CLI end-to-end, incl.\ Python startup) & --- & 0.45\,s \\
default (full plots, intermediates)           & ---     & 24.5\,s \\
\hline
\end{tabular}
\end{center}
The \texttt{--skip-plots --save-intermediates} number is the
relevant one for batched campaigns: it is now $\sim 280\times$
faster than the buggy baseline, and produces the same 29
\texttt{*\_FF.npy} spectra bit-for-bit (verified by reloading the
saved files and comparing
\texttt{max(|disk-mem|)}, which is $0$ to floating-point exact
on the test datasets).

\paragraph{Verification.}
The complex-FFT linearity used by \texttt{compute\_2d\_fft\_combo}
is exact in floating point (it does one extra complex-add per
spectrum, no extra rounding versus
$\lvert\mathcal F\{aX+bY\}\rvert$, which itself does $aX+bY$
followed by fft2).  The end-to-end check ran the optimised reader
on both the \texttt{nocouple\_ctrl} (no-coupling control,
SU(3) response identically zero by symmetry --- which the new
code reproduces correctly) and a coupled \texttt{quick\_test}
dataset, and confirmed
\texttt{disk == in-memory} bitwise for every $M_{NL}$,
$M_0$, $M_1$, $M_{01}$, and $\lambda$-mode spectrum.

\paragraph{Why this matters.}
Before XIX, an analyst sweeping over thirty pump-probe runs paid
$\sim 9\,$min just for reader-side work, mostly producing
debug PDFs that get overwritten on the next call.  After XIX,
the same campaign takes $\sim 14\,$s, which is short enough to
re-run on every parameter change.  The full-plot path is
preserved at 24\,s (slightly slower than 17\,s because the
$\lambda$-mode debug PDFs are now \emph{always} produced when
asked for, rather than the silently-skipped state we noted in
the bug above) and remains the right mode for one-off
exploratory analysis.

% =================================================================
\section{Ingredient XX: 2DCS pulse-chunking grid alignment and
envelope-window correctness}
% =================================================================

Ingredient~XV (W3) introduced \texttt{build\_pulse\_segments} so the
pump--probe and 2DCS drivers integrate the unperturbed dynamics with
a much larger \texttt{dt} hint outside the pulse-active window.
Ingredient~XVII later forced the observer cadence to be exactly
\texttt{T\_step} in every chunk, after a long-running rank-$K$
segfault was traced back to the chunked path producing trajectory
lengths that depended on the pulse positions.  The ``MPI 2DCS audit''
in this section closes two further sharp edges that survived
that fix.

\subsection*{Audit B1 -- segment endpoints must lie on the
\texttt{T\_step} grid}

Boost.Odeint's
\texttt{integrate\_const(controlled\_stepper,~system,~state,
$t_0$,~$t_1$,~$\Delta t$,~observer)} (the path used by every
controlled stepper -- \texttt{dopri5}, \texttt{rkck54},
\texttt{rkf78}, \texttt{bs}) advances the state by $\Delta t$
chunks while $\text{time} + \Delta t \le t_1$, fires a final
observer at the largest $t_0 + N\,\Delta t \le t_1$, and
\emph{drops the trailing remainder} $t_1 - (t_0 + N\,\Delta t)$
without integrating it.  The next chunk in the W3 driver then
re-enters \texttt{integrate\_const} with $t_\text{start} = t_1$
but the state in memory is still at $t_0 + N\,\Delta t$.  At every
seam where $(t_1 - t_0)$ is not an integer multiple of
\texttt{T\_step} the trajectory therefore loses up to \texttt{T\_step}
of integration time and silently propagates from a stale state.
For canonical TmFeO3 production runs (\texttt{T\_step}~=~0.01 with
\texttt{pump\_width}~=~0.4 or 0.5 and \texttt{tau\_step}~=~0.1) all
$\pm 6\sigma$ window endpoints, the merged window edges, and the
clamps at $T_\text{start},\,T_\text{end}$ are exact integer
multiples of \texttt{T\_step}, so the bug is invisible.  Any user
who picks
\texttt{pulse\_width}~=~0.45 or \texttt{tau\_step}~=~0.013 walks into
it without warning.

The fix is local to \texttt{include/classical\_spin/lattice/pulse\_chunking.h}:
\texttt{build\_pulse\_segments} now snaps every candidate window
endpoint and the global $T_\text{end}$ to the nearest
$T_\text{start} + k\,\texttt{T\_step}$ before the merge sort, and
asserts at the bottom of the function that every seam is contiguous
on the grid and every chunk length is an integer $\times$
\texttt{T\_step}.  The snap is symmetric and at most
$\texttt{T\_step}/2$, so the chunking still captures the requested
$\sim 6\sigma$ envelope, and for currently-shipped configs the
output of \texttt{build\_pulse\_segments} is bit-identical to the
old function.

A new pure-header regression test
\texttt{src/apps/test\_pulse\_chunking\_alignment.cpp} (registered
as \texttt{pulse\_chunking\_alignment} in CMake) exercises seven
cases -- canonical aligned, overlapping windows, misaligned
\texttt{pulse\_width}, misaligned \texttt{current\_tau}, misaligned
$T_\text{end}$, pulse outside window, and three degenerate inputs --
and verifies in each case that segments are on-grid, contiguous,
and that pulse centres stay inside an active region.

\subsection*{Audit B2 -- the documented $6\sigma$ window only
captures $e^{-9}$, not $10^{-16}$}

The drive envelope used by every \texttt{*\_pulse\_drive} RHS is
\[
  f(\Delta t) \;=\; A \, \exp\!\Bigl(-\bigl(\Delta t / (2\sigma)\bigr)^{2}\Bigr)
              \, \cos(\omega \, \Delta t)
\]
where \texttt{field\_drive\_width\_*} plays the role of $\sigma$ in
this expression.  Note the factor of two in the denominator: this
parameter is \emph{not} the standard Gaussian $\sigma$ but rather
$\sigma_\text{eff}/\sqrt{2}$.  At
$\Delta t = \texttt{kPulseWindowSigmas}\,\sigma$ the envelope
amplitude is
\[
  f(\Delta t) / A \;=\; \exp\!\bigl(-\texttt{kPulseWindowSigmas}^{2}/4\bigr).
\]
The pre-audit value $\texttt{kPulseWindowSigmas} = 6$ therefore gave
$\exp(-9) \approx 1.2 \times 10^{-4}$ -- four decimal places, not
the sixteen claimed by the surrounding header comment.  We bump
$\texttt{kPulseWindowSigmas}$ to $9$, giving
$\exp(-81/4) \approx 1.6 \times 10^{-9}$ -- nine decimal places,
genuinely below any practical 2DCS noise floor and consistent with
the original intent of the W3 hint.  The active window grows by
$50\%$, but only inside the (small) pulse-active regions, so the
free-segment $20\times$ \texttt{dt}-hint still dominates the W3
speedup; \texttt{test\_pulse\_drive\_tolerance} continues to pass at
its 5e-6 magnetisation ceiling.

\subsection*{Audit C1 -- symmetric Tukey window on the $\tau$ axis}

\texttt{util/readers\_new/reader\_TmFeO3.py} ships seven
apodization windows (Gaussian, exponential, Hann, Hamming, Blackman,
Tukey, cosine).  Six of them taper symmetrically around $\tau = 0$
via $|\tau|/\max|\tau|$.  The Tukey branch tapered only the $-\tau$
edge, leaking from the $+\tau$ side of any centred grid.  We mirror
the taper onto both edges so the window is symmetric and matches
the rest of the family.  This affected only users who chose
\texttt{window\_type='tukey'}; all other windows were already
correct.

\subsection*{Audit C2 -- clarifying the pump/probe cutoff comment}

The reader's $t > t_\text{cutoff}$ filter (\texttt{t\_cutoff
= pulse\_window\_sigma * pulse\_width}, default $5\sigma$) excludes
the \emph{first} pulse (always centred at $t = 0$ in the C++
convention used by \texttt{single\_pulse\_drive} and
\texttt{double\_pulse\_drive}).  The original comment described it
as ``skip the probe pulse region (probe is centered at t=0)'',
which is backwards from the standard 2DCS naming where the first
pulse is the pump.  More importantly the comment failed to mention
the \emph{second} pulse at $t = \tau$, which deliberately stays
inside the FFT window because that is exactly the
$\chi^{(3)}$ rephasing/non-rephasing signal we want to Fourier-resolve.
The math was already correct; we just rewrote the comment so the
next reader does not get confused into shifting the cutoff.

\subsection*{Things that look unusual but are correct}

\begin{itemize}[noitemsep]
\item \textbf{\texttt{\_flip\_omega\_tau}} (\texttt{reader\_TmFeO3.py}).
The reverse-on-axis-$-2$ operation implements the convention
$T \equiv -\tau$ for the displayed delay axis.  For an even
$n_\tau$ this is exact; for the canonical odd
$n_\tau = 2001$ ($\tau \in [-100, +100]$ at step $0.1$) it is off
by one bin in $\omega_\tau$, well below the resolution at which we
identify peaks.

\item \textbf{Boost \texttt{integrate\_const} drops the trailing
remainder of $[T_\text{start}, T_\text{end}]$} when
$(T_\text{end} - T_\text{start})/T_\text{step}$ is not an integer.
The audit B1 snap fixes this at the outer level too: $T_\text{end}$
is now snapped down to the largest grid point $\le T_\text{end}$
before any chunk is built.
\end{itemize}

% =================================================================
\section{Further reading}
% =================================================================
\begin{itemize}[noitemsep]
  \item D.~P.~Landau and K.~Binder, \emph{A Guide to Monte Carlo Simulations
        in Statistical Physics}, 4th ed., Cambridge UP (2014).
  \item U.~Wolff, ``Collective Monte Carlo updating for spin systems'',
        \emph{Phys.\ Rev.\ Lett.}\ \textbf{62}, 361 (1989).
  \item R.~H.~Swendsen and J.-S.~Wang, ``Nonuniversal critical dynamics
        in Monte Carlo simulations'', \emph{Phys.\ Rev.\ Lett.}\
        \textbf{58}, 86 (1987).
  \item K.~Hukushima and K.~Nemoto, ``Exchange Monte Carlo method and
        application to spin glass simulations'',
        \emph{J.~Phys.\ Soc.\ Jpn.}\ \textbf{65}, 1604 (1996).
  \item H.~Katzgraber, S.~Trebst, D.~Huse, M.~Troyer, ``Feedback-optimized
        parallel tempering Monte Carlo'', \emph{J.\ Stat.\ Mech.}\
        P03018 (2006).
  \item U.~Drepper, ``What every programmer should know about memory'',
        \texttt{lwn.net} (2007). \emph{Free, classic, and
        still essentially correct---the canonical reference for the
        cache-hierarchy facts above.}
  \item R.~F.~L.~Evans \emph{et al.}, ``Atomistic spin model simulations of
        magnetic nanomaterials'', \emph{J.~Phys.: Condens.\ Matter}
        \textbf{26}, 103202 (2014). \emph{The \textsc{Vampire} paper.}
\end{itemize}

\end{document}
